### 2020-10-11 latex/ks.tex
reorganize everything

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

\title{PDE: Smoothing}
\date{June 2020}

\begin{document}

\maketitle

\section{One way of thinking}

First order Taylor expansion: 
\begin{align}
  u(x_i) \approx u(x) + \frac{\partial u}{\partial x}(x_i - x) 
\end{align}
Then 
\begin{align}
   y_i &\approx  u(x) + \frac{\partial u}{\partial x}(x_i - x) + \epsilon_i \\ 
    \hat{u}(x) &= \sum_{i=1}^n y_i w_i \\
   &= \sum_{i=1}^n \left( u(x) + \frac{\partial u}{\partial x}(x_i - x) + \epsilon_i \right) w_i \\
   &= u(x) + \sum_{i=1}^n  w_i \epsilon_i +  \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  w_i(x_i - x) \\ 
   \hat{u}(x) - u(x) &= \sum_{i=1}^n  w_i \epsilon_i +  \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  w_i(x_i - x) \\ 
     E\left[ \left(\hat{u}(x) - u(x)\right)^2 \right] &=  E\left[ \left(w_1\epsilon_1+w_2\epsilon_2+...+w_n\epsilon_n \right)^2 \right] \\&+  E\left[ \left( \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  w_i(x_i - x)\right)^2 \right] \\ 
     E\left[ \left(\hat{u}(x_i) - u(x_i)\right)^2 \right] &=  \sigma^2\sum_{i=1}^n  w_i^2  +  \left[ \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  w_i(x_i - x) \right]^2 
\end{align}

\subsection{Linear}
We let

\begin{align}
    \phi_b (x) = \frac{1}{\sqrt{2 \pi b}}\exp(\frac{-x^2}{2b})
\end{align}

where b is the band width parameter. Then we have our MSE at $x_i$

\begin{align}
   MSE_j =  \sigma^2\sum_{i=1}^n  \phi_b^2(x_j-y_i)  +  \left[ \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  \phi_b(x_j-y_i)\cdot(x_j - y_i) \right]^2 
\end{align}

Now we replace summation with integration

\begin{align}
     MSE_j =  \sigma^2 \int_\omega  \phi_b^2(x_i-y)dy  +  \left[ \frac{\partial u}{\partial x}\cdot \int_\omega  \phi_b(x_j-y_i)\cdot(x_j - y)dy \right]^2 \\
     = \sigma^2 \sqrt{2\pi \cdot b}
\end{align}

\subsection{Quadratic}

We let

\begin{align}
    \phi_b (x) = \frac{1}{\sqrt{2 \pi b}}\exp(\frac{-x^2}{2b})
\end{align}

where b is the band width parameter. Then we have our MSE at $x_i$

\begin{align}
        \left(\hat{u}(x_i) - u(x_i)\right)  &=  \sigma^2\sum_{i=1}^n  w_i^2  + \left[ \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  w_i(x_i - x) \right] +  \left[ \frac{\partial^2 u}{2\partial x^2}\cdot \sum_{i=1}^n  w_i(x_i - x)^2 \right] + \sum_{i=1}^n \epsilon_i w_i \\
        E\left[ \left(\hat{u}(x_i) - u(x_i)\right)^2 \right] &= Var\left[ \left(\hat{u}(x_i) - u(x_i)\right) \right] + (E\left[ \left(\hat{u}(x_i) - u(x_i)\right) \right])^2 \\ &=\sigma^2\sum_{i=1}^n  w_i^2  + \left[\frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  w_i(x_i - x)  +   \frac{\partial^2 u}{2\partial x^2}\cdot \sum_{i=1}^n  w_i(x_i - x)^2 \right]^2 \\
   MSE_j =  \sigma^2\sum_{i=1}^n  \phi_b^2(x_j-y_i) & +  \left[ \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  \phi_b(x_j-y_i)\cdot(x_j - y_i) + \frac{\partial^2 u}{2\partial x^2}\cdot \sum_{i=1}^n  \phi_b(x_j-y_i)\cdot(x_j - y_i)^2 \right]^2
\end{align}

Now we replace summation with integration

\begin{align}
     MSE_j &=  \sigma^2 \int_\omega  \phi_b^2(x_i-y)dy  +  \left[ \frac{\partial u}{\partial x}\cdot \int_\omega  \phi_b(x_j-y)\cdot(x_j - y)dy + \frac{\partial^2 u}{2\partial x^2}\cdot \int_{\omega}  \phi_b(x_j-y)\cdot(x_j - y)^2 dy\right]^2\\
     &= \frac{\sigma^2}{2\sqrt{\pi \cdot b}}   + 0 + \left[ \frac{\partial^2 u}{2\partial x^2}\cdot \int_{\omega}  \phi_b(x_j-y)\cdot(x^2_j -2x_j\cdot y + y^2) dy\right]^2\\
     &=  \frac{\sigma^2}{2\sqrt{\pi \cdot b}}   + \left[ \frac{\partial^2 u}{2\partial x^2}\cdot \int_{\omega}  \phi_b(x_j-y)(x^2_j)dy - 2 \int_{\omega}  \phi_b(x_j-y)\cdot (x_j\cdot y)dy +  \int_{\omega}  \phi_b(x_j-y)\cdot y^2 dy\right]^2\\
     &=  \frac{\sigma^2}{2\sqrt{\pi \cdot b}}   + \left[\frac{\partial^2 u}{2\partial x^2}\cdot (x^2_j - 2x^2_j + b + x^2_j) \right]^2\\
     &=  \frac{\sigma^2}{2\sqrt{\pi \cdot b}}  + \left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2
\end{align}

Set $h(b) = \frac{\sigma^2}{2\sqrt{\pi}}\cdot \frac{1}{\sqrt{b}}+   \frac{1}{4}   u''^2\cdot b^2$, and take derivative, then set $h'(b) = 0$
\begin{align}
    h'(b)&= -\frac{\sigma^2}{4\sqrt{\pi}}\cdot b^{-\frac{3}{2}}+\frac{1}{2} u''^\cdot b = 0\\
    b &=\left(\frac{\sigma^2}{2\sqrt{\pi} u''^2}\right)^{\frac{2}{5}} 
\end{align}

\section{Another way of thinking}

Gaussian Kernel:

$$w_{ij}(x_j, g)  = \exp[- \frac{(x_i - x_j)^2}{2g^2}]$$

We have

$$MSE(x_j, g) = E[\hat{\mu}(x_j) - \mu(x_j)]^2$$

We will need to minimize

$$MSE(g) = \sum_{j=1}^n MSE(x_j, g)$$

We have

$$\hat{\mu}(x_j, i) = \mu(x_j) + \sum_{i = 1}^{n} w_{ij}(\mu'(x_j)(x_i-x_j) + \epsilon_i)  $$

Where $\epsilon_i$ follows $N(0, \sigma^2)$

Hence

$$MSE(x_j,g) = E[\sum_{i = 1}^{n} w_{ij}(\mu'(x_j)(x_i-x_j) + \epsilon_i) ]^2$$
$$=Var(\sum_{i = 1}^{n} w_{ij}(\mu'(x_j)(x_i-x_j) + \epsilon_i)) + (E[\sum_{i = 1}^{n} w_{ij}(\mu'(x_j)(x_i-x_j) + \epsilon_i) ])^2$$
$$=\sigma^2\sum_{i=1}^n w_{ij}^2 + [\sum_{i=1}^n w_{ij}\mu'(x_j)(x_i-x_j)]^2$$

\section{Question}

\begin{align*}
    Y_i &= g(x_i)+\epsilon_i\\ 
    MSE &= \sum_{i=1}^N(\hat{Y}_i- Y_i)^2 = \sum_{i=1}^N(\hat{Y}_i- g(x_i))^2\\ 
    MISE &= \int_1^5[\hat{g}(x) - g(x)]^2dx =\int_1^5[\hat{Y}_i - g(x)]^2dx =  \sum_{i=1}^N[\hat{Y}_i - g(x)]^2\cdot dx
\end{align*}

\end{document}

\end{document}

### 2020-10-11 shiny/HeatEqn.Rmd
reorganize everything

---
title: "Heat Equation"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eruptions, echo=FALSE}
library(shiny)
library(ggplot2)

# PDE
Laplacian <- function(y_k, dx){
  delta <- NULL
  len = length(y_k)
  # boundary condition
  delta[1] =(y_k[2]-y_k[1])/dx^2 
  delta[len] = (y_k[len-1]- y_k[len])/dx^2
  delta[2: (len-1)] <- diff(diff(y_k)/dx) /dx
  return(delta)
}
HeatEqn <- function(y0, dx, C, sd, dt, Nt, history = TRUE){
  yhat = list()
  yhat[[1]] = y0
  for (i in 1:(Nt+1)) {
    g = C* yhat[[i]]/sd
    yhat[[i+1]] = yhat[[i]] + g* Laplacian(yhat[[i]], dx) * dt
  }
  ifelse(history, return(yhat), return(yhat[[Nt+2]]))
}

###################################################################

# Shinny

ui <- fluidPage(

  # Application title
  #titlePanel("Data Generator"),

  # Sidebar with a slider input for number of bins 
  fluidRow(
    column(4, 
      textInput(inputId = "seed",
                label = "Choose set.seed",
                value = "1"),

      textInput(inputId = "equation", 
                label = "Underlying function", 
                value = "x^2"),

      sliderInput("xrange",
                  "Range of x:",
                  min = -10,
                  max = 10,
                  value = c(-1,1) )
    ),
    column(4,
      sliderInput(inputId = "nobs",
                  label = "Number of observations",
                  min = 2,
                  max = 300,
                  value = 20),

      sliderInput(inputId = "mean",
                  label = "Mean of noise",
                  min = -10,
                  max = 10,
                  value = 0.25,
                  step = 0.01),

      textInput(inputId = "sdtext", 
                label = "Standard deviation of noise",
                value = "0.1")
     ),
    column(4, 
      sliderInput(inputId = "dt",
                  label = "Set value for dt",
                  min = 0,
                  max = 0.0001,
                  value = 0.0001),

      sliderInput(inputId = "Nt",
                  label = "Set value for Nt",
                  min = 1,
                  max = 4000,
                  value = 200),
      #textInput(inputId = "g", 
      #          label = "Value of g",
      #          value = "x^2")
      sliderInput(inputId = "C",
                  label = "Set value for C",
                  min = 0.01,
                  max = 10,
                  value = 1)
     )
  ),

  fluidRow(
    column(6, 
           plotOutput("Plot1")),
    column(6, 
           plotOutput("Plot2")
           )
  )
)

###################################################################

server <- function(input, output){

  equation = function(x){ eval(parse(text = input$equation)) }

   sdtext = function(x){ eval(parse(text = input$sdtext)) }

  # diffusi = function(x){ eval(parse(text = input$g)) }
  C =  function(x){ eval(parse(text = input$C)) }

  xvalue = reactive({ seq(input$xrange[1], input$xrange[2], length.out = input$nobs) })

  noise = reactive({
    set.seed(input$seed)
    rnorm(input$nobs, mean = input$mean, sd = input$sdtext)})

  noise2 = reactive({
    set.seed(input$seed)
    rnorm(input$nobs, mean = input$mean, sd = sdtext(xvalue()))})

  yvalue = reactive({ equation(xvalue()) + noise2() })

  #  hqvalue = reactive({
  #     HeatEqn(yvalue(), 0.001, g = input$g, dt = input$dt, Nt= input$Nt, history = F)
  #})

  output$Plot1 <- renderPlot({
    x = xvalue()
    dx = x[2] - x[1]
    y = yvalue()
    y1 = equation(x)
    hq = HeatEqn(y, dx, input$C, noise2()+0.001, input$dt, input$Nt, history = F)

    df <- data.frame(cbind(x,y,y1,hq))
    ggplot(df) + 
      geom_point(aes(y = y, x = x, color = '1')) +
      geom_line(aes(y = y1, x = x, color = '2'), linetype = "twodash") + 
      geom_line(aes(y = hq, x = x, color = '3')) + 
      scale_color_discrete(name = "Y", labels = c("observations", "real value", "fitted value")) +
      ggtitle("One Dimensional Finite Difference Method")

  })

  output$Plot2 <- renderPlot({
    x = xvalue()
    dx = x[2] - x[1]
    y = yvalue()
    y1 = equation(x)
    hq = HeatEqn(y, dx, input$C, noise2()+0.001, input$dt, input$Nt, history = F)
    ratio = y1 /(noise2()+0.001)
    g = ratio * input$C
    df2 <- data.frame(cbind(x,ratio,g))
    ggplot(df2) + 
      geom_line(aes(y = ratio, x = x, color = '1'), linetype = "twodash") + 
      geom_line(aes(y = g, x = x, color = '2')) + 
      scale_color_discrete(name = "Y", labels = c("ratio", "g")) +
      ggtitle("Smoothing Strategy: Tuning of g")

  })

}

shinyApp(ui = ui, server = server)
```

### 2020-10-11 shiny/ks.Rmd
reorganize everything

---
title: "Kernel Smoothing"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eruptions, echo=FALSE}
library(shiny)
library(ggplot2)

# PDE
gauss.kernel = function(x, y,ind,b,win,proportion){
  b = max(b, 1e-6)
  ran = (max(range(x)) - min(range(x)))*proportion

  xfix = x[ind]

  ub =  min(xfix + win*b, xfix + ran/2) #upperbound
  lb =  max(xfix - win*b, xfix - ran/2) #lowerbound

  xnew = x[x >= lb & x <= ub]
  ynew = y[x >= lb & x <= ub]
  k = exp(-((xfix - xnew)^2)/(2*b))
  yhat = sum(k*ynew)/sum(k)
  return(yhat)
}

gaussian.smooth = function(x,y,b,win,proportion){
  yhat = vector(length = length(y))
  for(i in 1:length(y)){
    yhat[i] = gauss.kernel(x,y,i,b[i],win,proportion)  # adaptive bandwidth
  }
  return(yhat)
}

###################################################################

# Shinny

ui <- fluidPage(

  # Application title
  #titlePanel("Kernel Smoother"),

  # Sidebar with a slider input for number of bins 
  fluidRow(
    column(4, 
           textInput(inputId = "seed",
                     label = "Choose set.seed",
                     value = "1"),

           textInput(inputId = "equation", 
                     label = "Underlying function", 
                     value = "x^4"),

           textInput(inputId = "u2", 
                     label = "Second Derivative of Underlying function", 
                     value = "12*x^2")
    ),
    column(4,
           sliderInput(inputId = "nobs",
                       label = "Number of observations",
                       min = 2,
                       max = 300,
                       value = 200),

           sliderInput("xrange",
                       "Range of x:",
                       min = -10,
                       max = 10,
                       value = c(-5,5) ),

           textInput(inputId = "sdtext", 
                     label = "Standard deviation of noise",
                     value = "100")
    ),
    column(4, 
           sliderInput(inputId = "win",
                       label = "Set Window",
                       min = 0,
                       max = 20,
                       value = 5),

           sliderInput(inputId = "proportion",
                       label = "Set Proportion of Range",
                       min = 1,
                       max = 10,
                       value = 5),

           sliderInput(inputId = "sim",
                       label = "Simulation Number",
                       min = 100,
                       max = 1000,
                       value = 100)
    )
  ),

  fluidRow(
    column(6, 
           plotOutput("Plot1")),
    column(6, 
           plotOutput("Plot2"))
  )
)

###################################################################

server <- function(input, output){

  equation = function(x){ eval(parse(text = input$equation)) }
  u2 = function(x){ eval(parse(text = input$u2)) }

  sdtext = function(x){ eval(parse(text = input$sdtext)) }

  xvalue = reactive({ seq(input$xrange[1], input$xrange[2], length.out = input$nobs) })

  noise = reactive({
    set.seed(input$seed)
    rnorm(input$nobs, mean = 0, sd = sdtext(xvalue()))})

  yvalue = reactive({ equation(xvalue()) + noise() })

  bvalue = reactive({(sdtext(xvalue())^2/(2*sqrt(pi)*u2(xvalue())^2))^(2/5) })

  output$Plot1 <- renderPlot({
    x = xvalue()
    y = yvalue()
    b = bvalue()
    u = equation(x)
    ks = gaussian.smooth(x, y, b, input$win, input$proportion)

    plot(x,y, main= "Standard Deviation 100",pch = 20)
    lines(x, u, col = "blue")
    lines(x, ks, col = "red")  
    legend("top", legend = c("Underlying curve", "Optimal Curve"), col = c("blue", "red"), bty = "n", lty = 1)

  })

  output$Plot2 <- renderPlot({
    x = xvalue()
    y = yvalue()
    b = bvalue()
    u = equation(x)
    ks = gaussian.smooth(x, y, b, input$win, input$proportion)

    SSE = (gaussian.smooth(x,y,b,input$win, input$proportion)- u)^2     # remvoe the first and last 5 points
    MSE = mean(SSE, na.rm = T)  # remvoe the NA point (when x=0)

    bandw = 10^(seq(-8,8,0.5))
    MSE_b = NULL
    for (i in 1:length(bandw)) {
      SSE_vec  = (gaussian.smooth(x,y,rep(bandw[i], length(x)), input$win, input$proportion) - u)^2
      MSE_b[i] = mean(SSE_vec)
    }

    plot(seq(-8,8,0.5), MSE_b,pch = 20, ylab = "MSE", xlab = "bandwidth (10^x)", ylim = c(0, max(MSE_b)), xlim = c(-8,8), main = "MSE for different bandwidth value" )
    abline(h = MSE, col = "red") #MSE of optimal bandwidth
    legend("topleft", legend = c("Optimal bandwidth"), col = c("red"), lty = 1,  bty = "n")

  })

}

shinyApp(ui = ui, server = server)

```

### 2020-10-11 oldFiles/HeatEqn.v1.R

@@ -44,3 +44,19 @@ ggplot(df) +
   ggtitle("One Dimensional Finite Difference Method")
 
 
+#=====================================================================================
+# Comparison with Kernel Smoothing
+# Example 
+set.seed(1)
+x <- seq(-1,1,0.1)
+y <- x^2 + rnorm(length(x), 0, 0.1) # real value  
+hq <- HeatEqn(y0 = y, dx = 0.1, g = 1, dt = 0.001, Nt = 275, history = F)
+tail(hq)
+plot(x, y)
+plot(x, hq)
+plot(ksmooth(x, y, "normal",  bandwidth = 2))
+
+
+
+
+

### 2020-10-11 oldFiles/ks.v1.R

@@ -16,7 +16,7 @@ yhat = kernsmooth(x,y,2)
 plot(x,yhat, col=2)             
                
 #--Relationship between the ksmooth and finite difference--- 
-       2T = sigma^2  where T = dt*Nt and sigma = band*0.3706506    
+#   2T = sigma^2  where T = dt*Nt and sigma = band*0.3706506    
 
 #--------------------- Old Version --------------------------
 

### 2020-10-11 oldFiles/ks.v3.r

@@ -1,3 +1,6 @@
+#This code deals with the issue of boundary point. 
+#Note: this method works well only with seasonal data (aka. the data with sine or cosine curve)
+
 gauss.kernel = function(x, y,ind,b){
   dv = pmin(abs(x[ind] - x), x[length(x)]- x[1] - abs(x[ind] - x))  # circular boundary (pick up the minimum distance)
   k = exp(-(dv^2)/(2*b))*(1/sqrt(2*pi*b))

### 2020-10-11 oldFiles/ks.v4.R

@@ -1,4 +1,5 @@
-# limit finite window of kernel smoother
+# This code deals with the issue of bandwidth -> infinity. 
+# We set limit finite [Window] of the kernel smoother
 
 gauss.kernel = function(x, y,ind,b,win,proportion){
   b = max(b, 1e-6)

### 2020-10-11 source/gaussianCircular.R

@@ -1,117 +1,0 @@
-
-
-circular.dis = function(x1, xn, x, const){
-  start_point = x1 
-  end_point = xn
-  if(x <= const){
-    d1 = const - x
-    d2 = (x - start_point) + (end_point - const)
-    
-    return(min(d1^2, d2^2))
-  }else
-    d1 = const - x
-    d2 = (const - start_point) + (end_point - x)
-    return(min(d1^2, d2^2))
-}
-
-#testing the circular distance
-
-x = seq(1,11,1)
-circular.dis(1, 10, 1,10)
-
-#apply circular.dis for a vector
-circular.dis.vec = function(x, const){
-  y = rep(NA, length(x))
-  for (i in 1:length(x)) {
-    y[i] = circular.dis(x[1], x[length(x)],x[i],const)
-   
-  }
-  return(y)
-}
-
-#test circular.dis.vect
-
-z = circular.dis.vec(x,7)
-
-gauss.kernel = function(x, y,ind,b){
- # k = exp(-((x[ind] - x)^2)/(2*b))*(1/sqrt(2*pi*b))
-  k = exp(-circular.dis.vec(x,x[ind])/(2*b)) 
-  #print(k)
-  yhat = sum(k*y)/sum(k)
-  return(yhat)
-  
-}
-
-
-
-gaussian.smooth = function(x,y,b){
-  yhat = vector(length = length(y))
-  for(i in 1:length(y)){
-    yhat[i] = gauss.kernel(x,y,i,b[i])
-  }
-  return(yhat)
-}
-
-################################################################
-
-set.seed(1)
-x = seq(-2*pi + 1, 2*pi + 1, pi/32)
-u =  sin(x)
-
-sd = 1
-var = sd^2
-y = u + rnorm(length(u),0,sd)
-n = length(y)
-
-u2 = -sin(x)# #second derivative of u
-b = (var/(2*sqrt(pi)*u2^2))^(2/5) #calculate optimal bandwidth
-
-plot(x, b, ylim = c(-0,10))
-plot(x, b)
-plot(x, (u2)^2)
-
-
-
-plot(x,y, main= "Standard Deviation 100")
-lines(x, u, col = "blue")
-lines(x, gaussian.smooth(x, y, b), col = "red")
-lines(x,gaussian.smooth(x,y, rep(0.1, length(x))), col = "green")
-legend("top", legend = c("Underlying curve", "Optimal Curve"), col = c("blue", "red"), lty = 1)
-
[…]

### 2020-10-11 source/ks.simulation.r

@@ -1,9 +1,14 @@
+# simularion for 1000 times
+# plot MSE vs bandwidth of fix/optimal bandwidth
+# plot MSE vs x at every point
+
+##################### Function ##################################
+# Gaussian kernel without window 
 gauss.kernel2 = function(x, y,ind,b){
   k = exp(-((x[ind] - x)^2)/(2*b))*(1/sqrt(2*pi*b))
   yhat = sum(k*y)/sum(k)
   return(yhat)
 }
-
 gaussian.smooth2 = function(x,y,b){
   yhat = vector(length = length(y))
   for(i in 1:length(y)){
@@ -12,22 +17,19 @@ gaussian.smooth2 = function(x,y,b){
   return(yhat)
 }
 
+# Gaussian kernel with window 
 gauss.kernel = function(x, y,ind,b,win,proportion){
   b = max(b, 1e-6)
   ran = (max(range(x)) - min(range(x)))*proportion
-  
   xfix = x[ind]
-  
   ub =  min(xfix + win*b, xfix + ran/2) #upperbound
   lb =  max(xfix - win*b, xfix - ran/2) #lowerbound
-  
   xnew = x[x >= lb & x <= ub]
   ynew = y[x >= lb & x <= ub]
   k = exp(-((xfix - xnew)^2)/(2*b))
   yhat = sum(k*ynew)/sum(k)
   return(yhat)
 }
-
 gaussian.smooth = function(x,y,b,win,proportion){
   yhat = vector(length = length(y))
   for(i in 1:length(y)){
@@ -37,9 +39,7 @@ gaussian.smooth = function(x,y,b,win,proportion){
 }
 
 
-
-
-
+##################### Example ################################
 
 win = 2
 proportion = 0.2

### 2020-10-11 source/ksmooth.R

@@ -1,9 +1,0 @@
-# Example 
-set.seed(1)
-x <- seq(-1,1,0.1)
-y <- x^2 + rnorm(length(x), 0, 0.1) # real value  
-hq <- HeatEqn(y0 = y, dx = 0.1, g = 1, dt = 0.001, Nt = 275, history = F)
-tail(hq)
-plot(x, y)
-plot(x, hq)
-plot(ksmooth(x, y, "normal",  bandwidth = 2))

### 2020-10-11 source/ks.simulation.r
modified some comments

@@ -3,7 +3,7 @@
 # plot MSE vs x at every point
 
 ##################### Function ##################################
-# Gaussian kernel without window 
+# For fixed bandwidth (Gaussian kernel without window 
 gauss.kernel2 = function(x, y,ind,b){
   k = exp(-((x[ind] - x)^2)/(2*b))*(1/sqrt(2*pi*b))
   yhat = sum(k*y)/sum(k)
@@ -17,7 +17,7 @@ gaussian.smooth2 = function(x,y,b){
   return(yhat)
 }
 
-# Gaussian kernel with window 
+# For adaptive bandwidth (Gaussian kernel with window) 
 gauss.kernel = function(x, y,ind,b,win,proportion){
   b = max(b, 1e-6)
   ran = (max(range(x)) - min(range(x)))*proportion

### 2020-10-11 source/ks.simulation.r
modifies some comments

@@ -3,7 +3,7 @@
 # plot MSE vs x at every point
 
 ##################### Function ##################################
-# For fixed bandwidth (Gaussian kernel without window 
+# For fixed bandwidth (Gaussian kernel without window)
 gauss.kernel2 = function(x, y,ind,b){
   k = exp(-((x[ind] - x)^2)/(2*b))*(1/sqrt(2*pi*b))
   yhat = sum(k*y)/sum(k)

### 2020-10-21 src/Simulations_ver1.Rmd
Xing's first check-in. I moved your old files to the oldFiles directory, created ks.r to store all R functions, Simulations_ver1.Rmd for conducting simulations, and renamed the source directory to src, which conforms to the naming convention. I also created a file called .gitignore, so certain results files won't be tracked by github.

---
title: Simulations on Adaptive Kernel Smoothers
author: Qiuyi Wu, Cuong Pham, Xing Qiu
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: "Optional: An abstract with font Arial 10 to describe the study."
output:
  pdf_document:
    fig_height: 6
    fig_width: 6
    fig_caption: true
---

```{r init, echo=FALSE, include=FALSE, message=FALSE}
library(knitr)
library(rmarkdown)
library(pander)
## library(captioner)
## set some global options for the R code chunks
opts_chunk$set(echo=FALSE, message=FALSE, cache=TRUE, warning=FALSE, results='hide', fig.path='results/fig', dpi=300, cache.path='.cache/')

set.seed(2569)

####  Use the following line to manually generate the report
## render("Simulations_ver1.Rmd", output_format="pdf_document")

## To produce a manual newline between tables
newline <- function(n=1) { cat(paste(rep("&nbsp;\n", n), collapse=" ")) }

## load kernel smoothing functions
source("ks.r")
## Other commonly used packages
## library(lmerTest)

```

# Introduction

## Research aims

In this study, we aim to investigate the optimal adaptive smoothing method based on Gaussian kernel smoothing.

# Method

We implemented two kernel smoothers, one for Euclidean spaces and the other for $S^1$.

... more descriptions, including the choice of window size, our "bandwidth" is actually the standard deviation of the Gaussian kernel smoother, etc.

Assume that we know the oracle noise variance and the second order derivative of the function to be estimated, the optimal bandwidth

The optimal bandwidth is
$$ b^{*} = \left( \frac{\sigma^{2}(x)}{2\sqrt{\pi} (u''(x))^2} \right)^{2/5}.  $$

Mathematical derivations of the above formula is given in a separate document.

# Simulations

## Sim1: Estimating a Quadratic Function on $[-5, 5]$

```{r sim1-true-signal, fig.cap='Optimal bandwidth computed based on our theoretical derivations.'}

## specify parameters that generates data in Sim1
N <- 501
## Note that you shouldn't use "sd" as an object; it *replaces* an
## important system function sd()!!
sd1 <- 50; var1 <- sd1^2

## an evenly-spaced grid of x defined on [-5,5]
x <- seq(-5, 5, length.out=N); dx <- 10/(N-1)
## the true function to be estimated
u <- x^4+600
## its second order derivative
D2u <- 12*x^2
## the optimal bandwidth
b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)

## the fixed bandwidths
min.logb <- -1; max.logb <- -0.7; nb <- 11
logbs <- seq(min.logb, max.logb, length.out=nb)

## number of repetitions
reps <- 200

plot(x, b, type="l", ylim=c(0,5), xlab="x", ylab="Adaptive bandwidth")

```

In the first simulation, we aim to estimate the following function

$$ u(x) := x^4 + 600, \qquad x\in [-5, 5]. $$

To make our estimation easier, I decide to sample $u(x)$ on a relatively dense grid (a total number of N=`r N` evenly spaced points) on $[-5, 5]$. The observed data are the summation of $u(x_{i})$ and $i.i.d.$ normal noise $\epsilon_{j} \sim N(0, \sigma^2)$, $\sigma = `r sd1`$.

We compared the use of optimal adaptive bandwidth with fixed bandwidth (with $\log_{10} b$ ranges from `r min.logb` to `r max.logb`) in Gaussian kernel smoother.  The discrete data were smoothed by different choice of bandwidths and compared with the true underlying function $u(x)$. This simulation was repeated for $m=`r reps`$ times, and we use MSE averaged over all $x$ and repetitions to evaluate the performance of these methods.

```{r Sim1, fig.cap="Relationship between the smoothing bandwidth and mean MSE. The thick horizontal broken line on the bottom represents the mean MSE of the optimal adaptive kernel smoother; dots and thin solid curve represent the mean MSE obtained from kernel smoother with various fixed bandwidths; dotted curves represent +/- two standard errors from the mean MSEs. "}

MSEtab <- matrix(0, nrow=reps, ncol=nb+1)
colnames(MSEtab) <- c("Adaptive", paste0("logb=",logbs))

set.seed(123)
for (j in 1:reps){
  print(paste0("Repetition: ", j))
  y <- u + sd1*rnorm(N)
  yhat.adapt <- ks(x, y, bandwidth=b)
  yhat.fixed <- sapply(logbs, function(lb) ks(x,y,bandwidth=10^lb))
  Res <- sweep(cbind(yhat.adapt,yhat.fixed), 1, u)
  MSEtab[j,] <- colMeans(Res^2)
}

MSE.mean <- colMeans(MSEtab); MSE.se <- apply(MSEtab, 2, sd)/sqrt(reps)

plot(logbs, MSE.mean[-1], type="b", ylim=range(MSE.mean),
     xlab="log-bandwidth", ylab="Mean MSE")
## +/- 2*STDERR
lines(logbs,MSE.mean[-1]+2*MSE.se[-1], lty=3)
lines(logbs,MSE.mean[-1]-2*MSE.se[-1], lty=3)
abline(h=MSE.mean[1], lwd=2, lty=2)

```

From this figure, we see that the adaptive smoother had the smallest mean MSE, which is better than even the best fixed bandwidth smoother.

Finally, we take an arbitrary repetition (the last repetition) and show the quality of smoothing by the adaptive smoother and several choices of fixed bandwidths.

```{r Sim1.smoothing.plot, fig.cap="The true function and functions estimated by various smoothing methods. True underlying function is marked in black, adaptive smoothing in red, two choices of fixed bandwidth parameters are in green (small bandwidth, $\\log_{10} b=-1$) and blue (large bandwidth, $\\log_{10} b=-0.7$). The left panel shows all data points, and the right panel is a zoomed-in version which focuses on the right boundary. ", fig.width=12, fig.height=6}

par(mfrow=c(1,2))
plot(x, y, pch="."); lines(x,u)
lines(x, yhat.adapt, col=2)
lines(x, yhat.fixed[, 1], col=3)
lines(x, yhat.fixed[, 11], col=4)

plot(x, y, pch=1, xlim=c(4.2, 5), ylim=c(900, 1200)); lines(x,u)
lines(x, yhat.adapt, col=2)
lines(x, yhat.fixed[, 1], col=3)
lines(x, yhat.fixed[, 11], col=4)

```

As we can see from the left panel in this figure, adaptive smoothing had the best fit in the middle part of the figure; both green and blue curves *overfitted* (i.e., *under-smoothed*) the data, suggesting that a **larger** value of bandwidth should be used in that region.  However, judging from the right panel, the blue curve already started to *underfit* (over-smooth) the data at the boundary. It implies that using a larger bandwidth will only lead to worse fitting near the two boundaries. In short, the optimal adaptive smoothing achieved the best trade-off between fitting the middle (requires more smoothing) and the boundaries (requires less smoothing).

## Sim2: Estimating a Periodic Function on $S^1$

More work is needed.

```{r temp.area, eval=FALSE}

## temporary scripts

set.seed(123)
## xgrids with unequal steps
N <- 200; x <- sort(runif(N, min=-1, max=2))
## x <- seq(-1,1,.2); N <- length(x)
y <- x^2 + 3 + 0.5*rnorm(N)
## fixed bandwidth smoothing
yhat1 <- ks.fixed(x,y, .1, .3)
yhat2 <- ks.fixed(x,y, .5, 1)
yhat3 <- ks.fixed(x,y, 1,3)
## variable bandwidth smoothing
bb <- 10^seq(2,-3, length.out=N)
bb[1:5] <- Inf; bb[190:N] <- 0
## bb <- rep(Inf, N)
#bb <- seq(1, .1, length.out=N)
wins <- rep(1,N)
## yhat4 <- ks.variable(x,y, bb, wins)
yhat4 <- ks.variable(x,y, bb, wins)
## using the wrapper
yhat5 <- ks(x,y,bb)

plot(x,y)
lines(x,yhat1)
## points(x,yhat2,pch="x"); lines(x,yhat2, col=2)
lines(x,yhat3, col=3)
lines(x,yhat4, col=4)
lines(x,yhat5, col=5)

```

### 2020-10-21 src/ks.r

@@ -1,0 +1,188 @@
+## this file contains all the R functions required for a kernel
+## smoother. Do *NOT* mix simulations with functions!!
+
+## ks() is the main function implemented by Xing. Note that the main
+## purpose for writing this function is to demonstrate the underlying
+## mathematics, not for production-level use. For example, the current
+## function is based on the very inefficient for-loop in R; we almost
+## certainly will need to replace it by C/C++ in a production-level R
+## package. See ksmooth.c included in R's source code for a reference
+## implementation. Besides, it may be more efficient to implement KS
+## based on fast discrete Fourier transformation for fixed bandwidth.
+## Besides, for a production-level R package, we also need to add many
+## QC code to it, for example, it should stop if length(x) !=
+## length(y); it should have a mechanism to deal with NAs and probably
+## outliers; it also needs to check whether x is ordered or not -- KS
+## does not work for unsorted x!
+
+## x is the *ordered* vector of the covariate, y is the observed value
+## to be smoothed. bandwidth can either be a constant bandwidth or a
+## vector (adaptive bandwidth) that has the same length as x. In both
+## cases, the value (values) of bandwidth are defined as the standard
+## deviation of the Gaussian kernel. min.bandwith and max.bandwidth
+## are the minimum and maximum bandwidth. If bandwidth[i]
+## <min.bandwidth, we simply return y[j] as yhat[j]; if bandwidth[i] >
+## max.bandwidth, we use simple unweighted average over the smoothing
+## window as yhat[j]. By default ("auto"), min.bandwidth is set to be
+## min(xsteps)/4, where xsteps are the differences of two consecutive
+## x[j]; max.bandwidth is set to be 10*xrange, where xrange is the
+## range of the domain.
+
+## max.window is the maximum smoothing window, by default, it is set
+## to be 20% of xrange. For numerical reasons,
+## irrespective of the value of max.window, we don't need to extend
+## the smoothing window to more than 4*bandwidth.  Type is a
+## categorical variable that defines the type of kernel smoothing: its
+## default value "Euclidean" is the commonly used kernel smoother; the
[…]

### 2020-10-26 latex/ks.tex
Update ks.tex

@@
 \maketitle
+
+\section{Introduction}
+
+It is universally acknowledged that observed data in the physical world is, to some extent, contaminated with random noise.
+Hence, one can express our data generating mechanism using two components: a true underlying oracle function u(x) and a random component epsilon.
+It arises naturally that one would like to know u(x) as it is the data generator's essence.
+Yet, due to the fundamental sorrow -the limited observations of an infinite world, we fail to achieve the oracle function.
+Instead, we can only estimate it based on our finite noise infested observations.
+
+For decades, statisticians have been proposed various methods to estimate the oracle function u(x).
+One of the non-parametric ways is to use direct diffusion to smooth away all the noise.
+In an absolute non-rigorous sense, the direction diffusion process achieves the oracle risk by (weighted) averaging each observation neighbor.
+Each time we do this, our estimated function will be closer to the true function.
+Until, at a finite time T, we will achieve our goal.
+After that, at time T + 1, the estimated function will again be further away.
+People coin the term "over-smoothing" to indicate this phenomenon.
+Our goal is to discover the optimal T where we are closest to the truth given an observed data set.
+
+However, this seemingly simple task proves to be somewhat challenging.
+This is because the diffusion process is not widely used among statisticians.
+Luckily, it is known that the famous kernel (Gaussian) smoother and the direct diffusion (heat equation) are equivalents.
+Their connection is expressed through the equation T = 0.5 $\sigma$.
+T denotes the smoothing time in the direction process.
+And $\sigma$ denotes the spatial width of the Gaussian smoother.
+Therefore, finding optimal time T can be translated to the problem of finding the optimal sigma.
+In statistics literature, the optimal sigma is found to be $(1/2N\pi||f''||^2)^{2/5}$ a constant for every point x.
+This is a significant constraint.
+Hence, in this paper, we will allow the spatial bandwidth to be different for each point.
+In the next section, we will derive the optimal sigma for each observation point.
+We will do simulations to show that our "adaptive" bandwidth performs better than the best-fixed bandwidth.
 

### 2020-10-27 latex/draft.tex
update Latex draft

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{soul}
\usepackage{ulem}

\title{Kernel Smoothing And Heat Equation}
\date{\today}

\begin{document}

\maketitle

\section{Introduction}

It is universally acknowledged that observed data with respect to the underlying patterns we are seeking for in the physical world is, to some extent, contaminated with random noise. Commonly, the observed data would be expressed into two parts: the systematic component (i.e.: a true underlying oracle function $u(x)$) and a random component (i.e.: noise $\epsilon$). Yet, due to the fundamental sorrow of the limited observations in an infinite world, we may fail to achieve the oracle function. In fact, we can only estimate it based on our finite noise infested observations.

For decades, statisticians have been proposed various methods to estimate the oracle function. One of the non-parametric ways is to use direct diffusion to smooth away all the noise. In an absolute non-rigorous sense, the direction diffusion process achieves the oracle risk by (weighted) averaging each observation neighbor. Each time we do this, our estimated function will be closer to the true function until at a finite time $T$ we achieve our goal. After that the estimated function will be bounded away from the truth. People coin the term "over-smoothing" to indicate this phenomenon. Our goal is to discover the optimal T where we are closest to the truth given an observed data set.

However, the diffusion process is not widely used among statisticians. Instead, they use kernel smoothing, a nonparametric method proved to be equivalent to the heat equation in physics world. Therefore, finding optimal time T can be translated to the problem of finding the optimal sigma. 
In statistics literature (\ref{} reference needed), the optimal sigma is found to be a constant for every point x. This is a significant constraint. Hence, in this paper, we will allow the spatial bandwidth to be different for each point. In the next section, we will derive the optimal sigma for each observation point. We will do simulations to show that our "adaptive" bandwidth performs better than the best-fixed bandwidth.

Some literature review...

Goal of this paper...

Structure of the paper...

\section{Methodology}
The connection between Heat Equation and Kernel Smoothing can be expressed (\ref{}reference for Equation (\ref{eq:1})) as
\begin{equation}
    T = \frac{1}{2} \sigma^2         \label{eq:1}
\end{equation} 
where $T$ denotes the total smoothing time in the direction process, and $\sigma$ denotes the proportion of the spatial width of the Gaussian smoother in paper \ref{} (Equation (\ref{eq:2})).
\begin{equation}
    \sigma = \text{bandwidth}\times 0.3706506  \label{eq:2}
\end{equation}
Given $\mathbf{x}_i =(x_1, x_2,...,x_p)^\top \subseteq \mathbb{R}^p $, $Y_i\subseteq R$, we have the data set $$\mathcal{D}_n = \big\{(x_1, y_1),... (x_n, y_n)\big\}$$
Denote $u(\mathbf{x})$ as the oracle function, and $u(\mathbf{x}_i)$ is the first order Taylor expansion
\begin{equation}
      u(\mathbf{x}_i) \approx u(\mathbf{x}) + \frac{\partial u}{\partial \mathbf{x}}(\mathbf{x}_i - \mathbf{x}) 
\end{equation}
And given a Gaussian kernel with $b$ as the bandwidth parameter
\begin{align}
    \phi_b (x) = \frac{1}{\sqrt{2 \pi b}}\exp(\frac{-x^2}{2b})
\end{align}
We have MSE at $\mathbf{x}_i$ with respect to $b$
\begin{equation}
         \text{MSE}_j=  \frac{\sigma^2}{2\sqrt{\pi \cdot b}}  + \left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2
\end{equation}
Take the derivative with respective to $b$, we can solve the equation 
\begin{equation}
    b =\left(\frac{\sigma^2}{2\sqrt{\pi} u''^2}\right)^{\frac{2}{5}} 
\end{equation}
(Detailed calculation in Appendix)

%-------------- Draft --------------%

\pagebreak

\section{Appendix}

We let

\begin{align*}
    \phi_b (x) = \frac{1}{\sqrt{2 \pi b}}\exp(\frac{-x^2}{2b})
\end{align*}

where b is the band width parameter. Then we have our MSE at $x_i$

\begin{align*}
        \left(\hat{u}(x_i) - u(x_i)\right)  &=  \sigma^2\sum_{i=1}^n  w_i^2  + \left[ \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  w_i(x_i - x) \right] +  \left[ \frac{\partial^2 u}{2\partial x^2}\cdot \sum_{i=1}^n  w_i(x_i - x)^2 \right] + \sum_{i=1}^n \epsilon_i w_i \\
        E\left[ \left(\hat{u}(x_i) - u(x_i)\right)^2 \right] &= Var\left[ \left(\hat{u}(x_i) - u(x_i)\right) \right] + (E\left[ \left(\hat{u}(x_i) - u(x_i)\right) \right])^2 \\ &=\sigma^2\sum_{i=1}^n  w_i^2  + \left[\frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  w_i(x_i - x)  +   \frac{\partial^2 u}{2\partial x^2}\cdot \sum_{i=1}^n  w_i(x_i - x)^2 \right]^2 \\
   MSE_j =  \sigma^2\sum_{i=1}^n  \phi_b^2(x_j-y_i) & +  \left[ \frac{\partial u}{\partial x}\cdot \sum_{i=1}^n  \phi_b(x_j-y_i)\cdot(x_j - y_i) + \frac{\partial^2 u}{2\partial x^2}\cdot \sum_{i=1}^n  \phi_b(x_j-y_i)\cdot(x_j - y_i)^2 \right]^2
\end{align*}

Now we replace summation with integration

\begin{align*}
     MSE_j &=  \sigma^2 \int_\omega  \phi_b^2(x_i-y)dy  +  \left[ \frac{\partial u}{\partial x}\cdot \int_\omega  \phi_b(x_j-y)\cdot(x_j - y)dy + \frac{\partial^2 u}{2\partial x^2}\cdot \int_{\omega}  \phi_b(x_j-y)\cdot(x_j - y)^2 dy\right]^2\\
     &= \frac{\sigma^2}{2\sqrt{\pi \cdot b}}   + 0 + \left[ \frac{\partial^2 u}{2\partial x^2}\cdot \int_{\omega}  \phi_b(x_j-y)\cdot(x^2_j -2x_j\cdot y + y^2) dy\right]^2\\
     &=  \frac{\sigma^2}{2\sqrt{\pi \cdot b}}   + \left[ \frac{\partial^2 u}{2\partial x^2}\cdot \int_{\omega}  \phi_b(x_j-y)(x^2_j)dy - 2 \int_{\omega}  \phi_b(x_j-y)\cdot (x_j\cdot y)dy +  \int_{\omega}  \phi_b(x_j-y)\cdot y^2 dy\right]^2\\
     &=  \frac{\sigma^2}{2\sqrt{\pi \cdot b}}   + \left[\frac{\partial^2 u}{2\partial x^2}\cdot (x^2_j - 2x^2_j + b + x^2_j) \right]^2\\
     &=  \frac{\sigma^2}{2\sqrt{\pi \cdot b}}  + \left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2
\end{align*}

Set $h(b) = \frac{\sigma^2}{2\sqrt{\pi}}\cdot \frac{1}{\sqrt{b}}+   \frac{1}{4}   u''^2\cdot b^2$, and take derivative, then set $h'(b) = 0$
\begin{align*}
    h'(b)&= -\frac{\sigma^2}{4\sqrt{\pi}}\cdot b^{-\frac{3}{2}}+\frac{1}{2} u''^\cdot b = 0\\
    b &=\left(\frac{\sigma^2}{2\sqrt{\pi} u''^2}\right)^{\frac{2}{5}} 
\end{align*}

\end{document}

\end{document}

### 2020-10-27 src/ks.r
Fixed a bug in ks.r: an internal variable N was not defined in the ks.fixed() and ks.variable() functions.

@@ -43,7 +43,7 @@ ks.fixed <- function(x, y, bandwidth, smooth.window, type=c("Euclidean", "circul
   ## match.arg gives warning when the input is not one of the
   ## declared values
   type <- match.arg(type)
-  yhat <- rep(0, N)
+  N <- length(x); yhat <- rep(0, N)
   if (type=="Euclidean") {
     for (i in 1:N){
       ## y.within.window.R contains all values to be smoothed to the
@@ -99,7 +99,7 @@ ks.variable <- function(x, y, bandwidth, smooth.window, type=c("Euclidean", "cir
   ## match.arg gives warning when the input is not one of the
   ## declared values
   type <- match.arg(type)
-  yhat <- rep(0, N)
+  N <- length(x); yhat <- rep(0, N)
   if (type=="Euclidean") {
     for (i in 1:N){
       if (bandwidth[i]==0){

### 2020-10-27 latex/draft.tex
updated the draft.tex document.

@@
 \maketitle
+
+\section{Notes}
+\label{sec:notes}
+
+\begin{itemize}
+\item 10/27/2020. We reviewed Dr. Qiu's code/documentation. Todos:
+  \begin{enumerate}
+  \item Compare our old code with Dr. Qiu's new code one more time.
+  \item Try to implement the circular smoother (at least for the fixed bandwidth case).
+  \item We've already studied the mathematical connections between the classical (homogeneous) heat equation and Gaussian kernel smoothing.
+We need to document this important connection in this document.
+  \item We need to think: what is the mathematical connection between \emph{adaptive Gaussian kernel smoother} and \emph{inhomogeneous} heat equation?
+Rationale: we've already derived (based on large sample theory) the ``optimal'' variable bandwidth for kernel smoother.
+If the kernel smoother is equivalent to an inhomogeneous differential equation, it means that we can then use an efficient numerical DE solver (e.g., those based on finite element method) to do kernel smoothing, especially for multi-dimensional cases.
+  \item In the long run, we need to develop a practical estimation procedure for $x(t)$ and $\sigma^{2}(t)$, better with some model selection procedure so the entire estimation procedure can be automated.
+  \item Multi-dimensional!! (a) the curse of dimensionality, (b) much more complicated boundary to deal with.
+  \end{enumerate}
+\end{itemize}
 

### 2020-11-03 latex/draft.tex
updated meeting notes

@@
 \begin{itemize}
+% 11/03/2020  
+\item 11/03/2020.
+We implemented circular smoother on both fixed bandwidth and adaptive bandwidth scenarios; and did the corresponding simulation.
+We notice adaptive bandwidth cannot beat the optimal fixed bandwidth.
+Fixed bandwidth achieve the smallest MSE in our $y=\sin(x)$ setting.
+\begin{verbatim}
+           euclidean   circular
+Adaptive  0.19821445 0.19821445
+logb=-3   1.00313133 1.00313133
+logb=-2.4 1.00313133 1.00313133
+logb=-1.8 1.00313133 1.00313133
+logb=-1.2 0.11346202 0.11227971
+logb=-0.6 0.02875581 0.02717735
+logb=0    0.08004308 0.08145240
+logb=0.6  0.44585526 0.52684029
+logb=1.2  0.49326980 0.50454607
+logb=1.8  0.50018979 0.50094139
+logb=2.4  0.50068896 0.50068896
+logb=3    0.50068896 0.50068896
+\end{verbatim}
+
+TODOLIST:
+  \begin{enumerate}
+    \item 
+    \item 
+  \end{enumerate}
+
+% 10/27/2020  
 \item 10/27/2020. We reviewed Dr. Qiu's code/documentation. Todos:

### 2020-11-03 src/Simulations_ver1.Rmd

@@
 ---
-title: Simulations on Adaptive Kernel Smoothers
-author: Qiuyi Wu, Cuong Pham, Xing Qiu
+title: "Simulations on Adaptive Kernel Smoothers"
+author: "Qiuyi Wu, Cuong Pham, Xing Qiu"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
-abstract: "Optional: An abstract with font Arial 10 to describe the study."
 output:
   pdf_document:
+    fig_caption: yes
     fig_height: 6
     fig_width: 6
-    fig_caption: true
+  html_document:
+    df_print: paged
+abstract: 'Optional: An abstract with font Arial 10 to describe the study.'
 ---
@@
   y <- u + sd1*rnorm(N)
-  yhat.adapt <- ks(x, y, bandwidth=b)
-  yhat.fixed <- sapply(logbs, function(lb) ks(x,y,bandwidth=10^lb))
+  yhat.adapt <- ks(x, y, bandwidth=b, type = "Euclidean")
+  yhat.fixed <- sapply(logbs, function(lb) ks(x,y,bandwidth=10^lb, type = "Euclidean"))
   Res <- sweep(cbind(yhat.adapt,yhat.fixed), 1, u)
@@
 
-## Sim2: Estimating a Periodic Function on $S^1$
+## Sim2: Estimating a Periodic Function on $[-2\pi, 2\pi]$
+```{r sim2-true-signal, fig.cap='Optimal bandwidth computed based on our theoretical derivations.'}
 
-More work is needed.
+## specify parameters that generates data in Sim2
+N <- 501
+## Note that you shouldn't use "sd" as an object; it *replaces* an
+## important system function sd()!!
+sd1 <- 1; var1 <- sd1^2
+
+## an evenly-spaced grid of x defined on [-5,5]
+x <- seq(-2*pi, 2*pi, length.out=N); dx <- 10/(N-1)
+## the true function to be estimated
+u <- sin(x)
+## its second order derivative
+D2u <- cos(x)
+## the optimal bandwidth
+b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+
+## the fixed bandwidths
+min.logb <- -3; max.logb <- 3; nb <- 11
+logbs <- seq(min.logb, max.logb, length.out=nb)
+
+## number of repetitions
+reps <- 200
+
+plot(x, b, type="l", ylim=c(0,5), xlab="x", ylab="Adaptive bandwidth")
+
+```
+In this simulation, we aim to estimate the following function
+$$u(x) := \sin(x), \qquad x\in [-2\pi, 2\pi] $$
+
+Again we use the same grid (a total number of N=`r N` evenly spaced points) on $[-2\pi, 2\pi]$.
+The observed data are the summation of $u(x_{i})$ and $i.i.d.$ normal noise $\epsilon_{j} \sim N(0, \sigma^2)$, $\sigma = `r sd1`$.
+
+We compared the use of optimal adaptive bandwidth with fixed bandwidth (with $\log_{10} b$ ranges from `r min.logb` to `r max.logb`) in Gaussian kernel smoother.
+The discrete data were smoothed by different choice of bandwidths and compared with the true underlying function $u(x)$.
+This simulation was repeated for $m=`r reps`$ times, and we use MSE averaged over all $x$ and repetitions to evaluate the performance of these methods.
+
+```{r Sim2, fig.cap="Relationship between the smoothing bandwidth and mean MSE.
+The thick horizontal broken line on the bottom represents the mean MSE of the optimal adaptive kernel smoother; dots and thin solid curve represent the mean MSE obtained from kernel smoother with various fixed bandwidths; dotted curves represent +/- two standard errors from the mean MSEs.
+"}
+MSEtab <- matrix(0, nrow=reps, ncol=nb+1)
+colnames(MSEtab) <- c("Adaptive", paste0("logb=",logbs))
+set.seed(123)
+for (j in 1:reps){
+  print(paste0("Repetition: ", j))
+  y <- u + sd1*rnorm(N)
+  yhat.adapt <- ks(x, y, bandwidth=b, type = "circular")
+  yhat.fixed <- sapply(logbs, function(lb) ks(x,y,bandwidth=10^lb, type = "circular"))
+  Res <- sweep(cbind(yhat.adapt,yhat.fixed), 1, u)
+  MSEtab[j,] <- colMeans(Res^2)
+}
+
+MSE.mean <- colMeans(MSEtab); MSE.se <- apply(MSEtab, 2, sd)/sqrt(reps)
+plot(logbs, MSE.mean[-1], type="b", ylim=range(MSE.mean),
+     xlab="log-bandwidth", ylab="Mean MSE")
+## +/- 2*STDERR
+lines(logbs,MSE.mean[-1]+2*MSE.se[-1], lty=3)
+lines(logbs,MSE.mean[-1]-2*MSE.se[-1], lty=3)
+abline(h=MSE.mean[1], lwd=2, lty=2)
+```
+
+From this figure, we see that the adaptive smoother cannot beat the fixed bandwidth smoother. Fixed bandwidth smoother with `r names(which(MSE.mean==min(MSE.mean)))` achieved the smallest mean MSE.
+
+Finally, we take an arbitrary repetition (the last repetition) and show the quality of smoothing by the adaptive smoother and several choices of fixed bandwidths.
+
+```{r Sim2.smoothing.plot, fig.cap="The true function and functions estimated by various smoothing methods.
+True underlying function is marked in black, adaptive smoothing in red, two choices of fixed bandwidth parameters are in green (small bandwidth, $\\log_{10} b=-0.6$) and blue (the smallest fixed bandwidth, $\\log_{10} b=0$).
+The left panel shows all data points, and the right panel is a zoomed-in version which focuses on the right boundary.
+", fig.width=12, fig.height=6}
+
+plot(x, y, pch="."); lines(x,u)
+lines(x, yhat.adapt, col="red")
+lines(x, yhat.fixed[, 5], col="green")
+lines(x, yhat.fixed[, 6], col="blue")
+
+```
 

### 2020-11-10 latex/draft.tex
update latex file
Update notes after the reading class on Nov 10

@@
 \begin{itemize}
-% 11/03/2020  
-\item 11/03/2020.
+% 11/10/2020  
+\item 11/10/2020.
 We implemented circular smoother on both fixed bandwidth and adaptive bandwidth scenarios; and did the corresponding simulation.
-We notice adaptive bandwidth cannot beat the optimal fixed bandwidth.
+We noticed for the circular data, the adaptive bandwidth cannot beat the optimal fixed bandwidth.
 Fixed bandwidth achieve the smallest MSE in our $y=\sin(x)$ setting.
+We also noticed the circular smoother performs better than Euclidean smoother in the circular data ($y=\sin(x)$).
 \begin{verbatim}
@@
   \begin{enumerate}
-    \item 
-    \item 
+    \item Compare MSE at 2 boundaries 
+    \item manually do the smoothing for certain points, compare the two bandwidth (adaptive vs fixed); expected MSE and empirical MSE should agree. Two scenarios:
+    \begin{enumerate}
+    \item manual result doesn't equal to the automatic one, debug the result
+    \item manual result equal to the auto one. Seek the reason: Taylor expansion is not good enough, higher order term is needed (More observation, smaller bandwidth, smaller window)
+    \end{enumerate}
+    \item Future: explore the inhomogeneous heat equation (adaptive $\lambda(x)$) 
+    $$\frac{\partial f(x,t)}{\partial t}  = \Delta f = \frac{\partial f}{\partial x}[\lambda(x)\frac{\partial f}{\partial x}] \neq  \lambda(x)\frac{\partial^2 f}{
+    \partial x^2} $$
+
+    Google the connection between adaptive smoothing and inhomogeneous heat equation 
   \end{enumerate}

### 2020-11-18 src/ks.r
update for simulation

@@ -247,7 +247,7 @@ ks <- function(x, y, bandwidth, min.bandwidth="auto", max.bandwidth="auto", max.
   type <- match.arg(type)
   N <- length(x); xsteps <- diff(x); xrange <- max(x)-min(x)
   if (min.bandwidth=="auto") min.bandwidth <- min(xsteps)
-  if (max.bandwidth=="auto") max.bandwidth <- 10*xrange
+  if (max.bandwidth=="auto") max.bandwidth <- xrange*10  # initially 10*xrange
   if (max.window=="auto") max.window <- xrange/5  # initially xrange/5
   ## classify the smoothing strategies based on bandwidth
   if (length(bandwidth)==1){ #fixed bandwidth
@@ -258,7 +258,7 @@ ks <- function(x, y, bandwidth, min.bandwidth="auto", max.bandwidth="auto", max.
       ## just return the simple sample mean
       yhat <- rep(mean(y), N)
     } else { #this is the general case
-      smooth.window <- min(400*bandwidth, max.window)  # initially 4*bandwidth
+      smooth.window <- min(4*bandwidth, max.window) # initially 4*bandwidth
       yhat <- ks.fixed(x, y, bandwidth, smooth.window, type=type)
     }
   } else if (length(bandwidth)!=N){
@@ -266,8 +266,8 @@ ks <- function(x, y, bandwidth, min.bandwidth="auto", max.bandwidth="auto", max.
   } else {
     ## process bandwith and smooth.window for ks.variable()
     bandwidth <- replace(bandwidth, bandwidth<=min.bandwidth, 0)
-    bandwidth <- replace(bandwidth, bandwidth>=max.bandwidth, Inf)
-    smooth.window <- pmin(400*bandwidth, max.window)  # initially 4*bandwidth
+    bandwidth <- replace(bandwidth, bandwidth>=max.bandwidth, Inf) 
+    smooth.window <- pmin(4*bandwidth, max.window) # initially 4*bandwidth
     yhat <- ks.variable(x, y, bandwidth, smooth.window)
   }
   return(yhat)

### 2020-11-18 src/simulation.R

@@ -1,0 +1,175 @@
+set.seed(2569)
+
+### Sim1: Estimating a Quadratic Function 
+
+## specify parameters that generates data in Sim1
+N <- 501
+## Note that you shouldn't use "sd" as an object; it *replaces* an
+## important system function sd()!!
+sd1 <- 50; var1 <- sd1^2
+
+## an evenly-spaced grid of x defined on [-5,5]
+x <- seq(-5, 5, length.out=N); dx <- 10/(N-1)
+## the true function to be estimated
+u <- x^4+600
+## its second order derivative
+D2u <- 12*x^2
+## the optimal bandwidth
+b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+
+## the fixed bandwidths
+min.logb <- -1; max.logb <- -0.7; nb <- 11
+logbs <- seq(min.logb, max.logb, length.out=nb)
+
+## number of repetitions
+reps <- 50
+plot(x, b, type="l", ylim=c(0,5), xlab="x", ylab="Adaptive bandwidth")
+
+# Relationship between the smoothing bandwidth and mean MSE
+MSEtab <- matrix(0, nrow=reps, ncol=nb+1)
+colnames(MSEtab) <- c("Adaptive", paste0("logb=",logbs))
+
+set.seed(123)
+for (j in 1:reps){
+  print(paste0("Repetition: ", j))
+  y <- u + sd1*rnorm(N)
+  yhat.adapt <- ks(x, y, bandwidth=b, type = "Euclidean")
+  yhat.fixed <- sapply(logbs, function(lb) ks(x,y,bandwidth=10^lb, type = "Euclidean"))
+  Res <- sweep(cbind(yhat.adapt,yhat.fixed), 1, u)
+  MSEtab[j,] <- colMeans(Res^2)
+}
+
+MSE.mean <- colMeans(MSEtab); MSE.se <- apply(MSEtab, 2, sd)/sqrt(reps)
+
+plot(logbs, MSE.mean[-1], type="b", ylim=range(MSE.mean),
+     xlab="log-bandwidth", ylab="Mean MSE")
+## +/- 2*STDERR
+lines(logbs,MSE.mean[-1]+2*MSE.se[-1], lty=3)
+lines(logbs,MSE.mean[-1]-2*MSE.se[-1], lty=3)
+abline(h=MSE.mean[1], lwd=2, lty=2)
+
+
+#par(mfrow=c(1,2))
[…]

### 2020-11-18 latex/draft.tex

@@
 \usepackage{ulem}
+\usepackage{graphicx}
+\usepackage{grffile} % allow use .eps files
+\usepackage{float}
 
@@
 \begin{itemize}
+% 12/01/2020  
+\item 12/01/2020. 
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pic/sim.plot1.eps}
+\includegraphics[width=\linewidth]{pic/sim.plot2.eps}
+\includegraphics[width=\linewidth]{pic/sim.plot3.eps}
+\end{figure}
+
+% 11/17/2020  
+\item 11/17/2020.
+
+TODOLIST:
+\begin{enumerate}
+\item Editor: Emacs, Vim
+\item Learn: Command line, \texttt{ssh}, \texttt{zsh},\texttt{bash}, to remotely control.
+\item \texttt{.eps} $\rightarrow$ \texttt{.eps}; change \texttt{postscript()} to \texttt{pdf(), png(), tiff()}
+\item Residual trend means under-smoothing (overfitting)
+\item manually compute the 2 extreme points (second order taylor expansion) -- empirical MSE and theoretical MSE
+\end{enumerate}
+
 % 11/10/2020  
@@
 We also noticed the circular smoother performs better than Euclidean smoother in the circular data ($y=\sin(x)$).
+
 \begin{verbatim}
@@
 \end{document}
-
-\end{document}

### 2020-12-01 latex/draft.tex
update latex file "

@@
 \includegraphics[width=\linewidth]{pic/sim.plot3.eps}
+\end{figure}
+
+The boundary behavior of the adaptive scenario is not ideal. We then modified $\texttt{max.window}$ from $\texttt{xrange/5}$ to $\texttt{xrange/0.1}$. The situation gets better.
+
+We explored the extreme points by 
+\begin{enumerate}
+\item computing the difference between the empirical value \texttt{mean(y.within.window)} and the true value \texttt{u}
+\item comparing the empirical MSE with the theoretical MSE of the extreme points
+\end{enumerate}
+
+Denote $\texttt{max.window} = \frac{\texttt{xrange}}{k}$, we plot the difference between the empirical value \texttt{mean(y.within.window)} and the true value \texttt{u} with respect to $k$ at extreme points.
+We noticed when $k>2$, the difference would increase tremendously (Figure below).
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pic/sim.plot4.eps}
+\includegraphics[width=\linewidth]{pic/sim.plot5.eps}
+\caption{Extreme point (first point) of $y=\sin(x)$ with $\sigma = 1$ and  $\sigma = 0.1$ scenarios.
+We use log-scale plot to see if the difference would converge asymptomatically as $k\rightarrow +\infty$.
+The red line denotes when $k=2$, the difference has a jump.}
 \end{figure}

### 2020-12-01 latex/draft.tex
update latex and code for debugging the extreme boundary point

@@
 
-The boundary behavior of the adaptive scenario is not ideal. We then modified $\texttt{max.window}$ from $\texttt{xrange/5}$ to $\texttt{xrange/0.1}$. The situation gets better.
+Although the adaptive bandwidth smoother perform well in general, but the boundary behavior of the adaptive bandwidth smoother is not ideal.
+We then modified $\texttt{max.window}$ from $\texttt{xrange/5}$ to $\texttt{xrange/0.1}$.
+The situation gets better.
 
@@
 The red line denotes when $k=2$, the difference has a jump.}
+\end{figure}
+
+Now we compare the empirical MSE with the theoretical MSE on the extreme point at the boundary.
+
+\textit{Flash back:}
+$$\text{MSE}_j=  \frac{\sigma^2}{2\sqrt{\pi \cdot b}}  + \left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2$$
+
+\begin{table}[H]
+\centering
+\begin{tabular}{l|l| l}
+\hline\hline
+& (theoretical MSE =$1.96\times 10^{-7}$) & (theoretical MSE =$3.95\times 10^{-7}$)\\
+k & empirical MSE  ($\sigma=1$) &empirical MSE  ($\sigma=0.1$) \\\hline 
+0.1&0.0019 $(8.15\times 10^{-6})$ & 0.0021 $(7.59\times 10^{-6})$\\
+1 & 0.0020 $(7.80\times 10^{-6})$ &0.0019 $(6.76\times 10^{-6})$\\
+2 & 0.0020 $(8.09\times 10^{-6})$ &0.0019 $(7.09\times 10^{-6})$\\
+2.1 & 0.4448 $(0.0072)$ &0.4408 (0.0071)\\
+3 & 0.5124 $(0.0128)$ & 0.5220 (0.0123)\\
+4 & 0.4107 $(0.0122)$ & 0.4070 (0.0131)\\
+5 & 0.3116 $(0.0124)$ & 0.3147 (0.0136)\\\hline
+\end{tabular}
+\caption{1000 times simulation for empirical MSE at extreme point at boundary}
+\end{table}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pic/sim.plot6.eps}
+\includegraphics[width=\linewidth]{pic/sim.plot7.eps}
+\includegraphics[width=\linewidth]{pic/sim.plot8.eps}
+\includegraphics[width=\linewidth]{pic/sim.plot9.eps}
 \end{figure}

### 2020-12-01 oldFiles/debug.R

@@ -1,0 +1,94 @@
+simk <- function(k,y){
+  smooth.window = 2*pi/k
+  y.within.window.R <- c(); dist.R <- c()
+  j <- 0; dj <- 0
+  while (dj<=smooth.window & j<=N-i) {
+    dj <- pmin(abs(x[i+j]-x[i]), x[length(x)]- x[1] - abs(x[i+j]-x[i])) 
+    dist.R <- c(dist.R, dj)
+    y.within.window.R <- c(y.within.window.R, y[i+j])
+    j <- j+1
+  }
+  
+  if (dj>smooth.window) {
+    y.within.window.R <- y.within.window.R[-j]
+    dist.R <- dist.R[-j]
+  }
+  
+  dd = dist.R
+  y.within.window = y.within.window.R
+  dnorm(dd, sd=b[1])
+  
+  sse = (mean(y.within.window)- u[1])^2
+  return(sse)
+  #return(mean(y.within.window)- u[1])
+}
+
+
+k=5
+SE = NULL
+for (m in 1:1000) {
+  y <- u + sd1*rnorm(N)
+  SE[m] = simk(k,y)
+}
+(MSE = mean(SE))
+(VAR = var(SE))
+(D2u[1]*b[1]/2)^2+sd1^2/(2*sqrt(pi*b[1]))
+
+(T.MSE = (D2u[126]*b/2)^2+sd1^2/(2*sqrt(pi*b)))
+
+
+
+
+set.seed(123)
+y <- u + 1*rnorm(N)
+
+
+filename = paste('sim.plot4.eps')
+setEPS()
+postscript(filename,height=6,width=16)
+layout(matrix(1:2,1,2, byrow = T))
+
+plot(seq(0.1,20, 0.05),unlist(lapply(seq(0.1,20, 0.05), simk, y=y)),pch =20, 
+     xlab = "k", ylab = "difference between Yhat and u", main = "sd=1")
+abline(v = 2, col=2, lwd = 3)
+plot(seq(-3,8, 1),unlist(lapply(10^(seq(-3,8, 1)), simk, y=y)),pch = 20,
+     xlab = "10^k", ylab = "difference between Yhat and u", main = "sd=1, log k scale")
+abline(v = log(2), col=2,lwd = 3)
+
+dev.off()
+
+
+
+
+set.seed(123)
+y <- u + 0.1*rnorm(N)
+
+filename = paste('sim.plot5.eps')
+setEPS()
[…]

### 2020-12-01 src/ks.r

@@ -248,7 +248,7 @@ ks <- function(x, y, bandwidth, min.bandwidth="auto", max.bandwidth="auto", max.
   N <- length(x); xsteps <- diff(x); xrange <- max(x)-min(x)
   if (min.bandwidth=="auto") min.bandwidth <- min(xsteps)
   if (max.bandwidth=="auto") max.bandwidth <- xrange*10  # initially 10*xrange
-  if (max.window=="auto") max.window <- xrange/5  # initially xrange/5
+  if (max.window=="auto") max.window <- xrange/0.1  # initially xrange/5
   ## classify the smoothing strategies based on bandwidth
   if (length(bandwidth)==1){ #fixed bandwidth
     if (bandwidth <= min.bandwidth) {

### 2020-12-01 src/simulation.R

@@ -137,7 +137,7 @@ legend("topleft", c("true function","Adaptive","logb=-0.6 (Opt)","logb=0"), col 
 
 #### save plots #########
 
-filename = paste('sim.plot3.eps')
+filename = paste('sim.plot8.eps')
 setEPS()
 postscript(filename,height=3,width=10)
 layout(matrix(1:3,1,3, byrow = T))
@@ -145,14 +145,15 @@ layout(matrix(1:3,1,3, byrow = T))
 
 MSE.mean <- colMeans(MSEtab); MSE.se <- apply(MSEtab, 2, sd)/sqrt(reps)
 plot(logbs, MSE.mean[-1], type="b", ylim=range(MSE.mean),
-     xlab="log-bandwidth", ylab="Mean MSE", main = "max.bandwidth = xrange*10, sd = 0.1")
+     xlab="log-bandwidth", ylab="Mean MSE", main = "max.bandwidth = xrange/0.1")
 ## +/- 2*STDERR
 lines(logbs,MSE.mean[-1]+2*MSE.se[-1], lty=3)
 lines(logbs,MSE.mean[-1]-2*MSE.se[-1], lty=3)
 abline(h=MSE.mean[1], lwd=2, lty=2)
 
 
-plot(x,Res[,1], ylim = c(-1,1),type = "l", ylab = "residual",main = "Resudial", col = 2)
+plot(x,Res[,1], ylim = c(-1,1),type = "l", ylab = "residual",
+     main = expression(paste("Resudial Plot, ",sigma," = 1, range = [",-pi,",",pi,"]")), col = 2)
 lines(x,Res[,6],  col = 3)
 lines(x,Res[,7],  col = 4)
 legend("top", c("Adaptive","logb=-0.6 (Opt)", "logb=0"), col =2:4, pch = 20 )

### 2020-12-08 latex/draft.tex
fixed boundary issue for circular one, and updated the notes  draft.

@@
 \begin{itemize}
+  % 12/08/2020  
+\item 12/08/2020. 
+Fixed the boundary issue for circular situation.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{l|l| l}
+\hline\hline
+& (theoretical MSE =$1.96\times 10^{-7}$) & (theoretical MSE =$4.93\times 10^{-9}$)\\
+k & empirical MSE  ($\sigma=1$) &empirical MSE  ($\sigma=0.1$) \\\hline 
+0.1&0.0020 $(6.65\times 10^{-6})$ & $2.02\times 10^{-5} (7.35\times 10^{-10})$\\
+1 & 0.0021 $(9.69\times 10^{-6})$ & $1.95\times 10^{-5}  (7.00\times 10^{-10})$\\
+2 & 0.0021 $(8.33\times 10^{-6})$ & $1.95\times 10^{-5} (7.17\times 10^{-10})$\\
+2.4 & 0.0024 $(1.22\times 10^{-5})$ & $2.41\times 10^{-5} (1.17\times 10^{-9})$\\
+3 & 0.0029 $(1.51\times 10^{-5})$ & $2.82\times 10^{-5} (1.59\times 10^{-9})$\\
+4 & 0.0041 $(3.52\times 10^{-5})$ &  $4.36\times 10^{-5} (3.42\times 10^{-9})$\\
+5 & 0.0050 $(5.36\times 10^{-5})$ &  $5.07\times 10^{-5} (5.24\times 10^{-9})$\\\hline
+\end{tabular}
+\caption{1000 times simulation for empirical MSE at extreme point at boundary}
+\end{table}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pic/sim.plot10.eps}
+\end{figure}
+\begin{verbatim}
+  Adaptive    logb=-3  logb=-2.4  logb=-1.8  logb=-1.2  logb=-0.6 
+0.01650001 1.00313133 1.00313133 0.22355199 0.05497094 0.01329125 
+    logb=0   logb=0.6   logb=1.2   logb=1.8   logb=2.4     logb=3 
+0.02305490 0.03210026 0.03280336 0.50068896 0.50068896 0.50068896
+\end{verbatim}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pic/sim.plot11.eps}
+\end{figure}
+\begin{verbatim}
+    Adaptive      logb=-3    logb=-2.4    logb=-1.8    logb=-1.2    logb=-0.6 
+0.0004494958 0.0100313133 0.0100313133 0.0022356024 0.0005502092 0.0005894302 
+      logb=0     logb=0.6     logb=1.2     logb=1.8     logb=2.4       logb=3 
+0.0197699286 0.0292714857 0.0299909440 0.4990188656 0.4990188656 0.4990188656 
+\end{verbatim}
+
+TODOLIST:
+\begin{enumerate}
+\item 
+\item
+\end{enumerate}
+
 % 12/01/2020  

### 2020-12-08 oldFiles/debug.R

@@ -1,26 +1,14 @@
 simk <- function(k,y){
   smooth.window = 2*pi/k
-  y.within.window.R <- c(); dist.R <- c()
-  j <- 0; dj <- 0
-  while (dj<=smooth.window & j<=N-i) {
-    dj <- pmin(abs(x[i+j]-x[i]), x[length(x)]- x[1] - abs(x[i+j]-x[i])) 
-    dist.R <- c(dist.R, dj)
-    y.within.window.R <- c(y.within.window.R, y[i+j])
-    j <- j+1
-  }
   
-  if (dj>smooth.window) {
-    y.within.window.R <- y.within.window.R[-j]
-    dist.R <- dist.R[-j]
-  }
+  dj <- pmin(abs(x-x[1]), x[length(x)]- x[1] - abs(x-x[1]))
+  y.within.window <- y[dj <= smooth.window]
+  #dd = dj[dj <= smooth.window]
+  #w <- dnorm(dd, sd=b[1])
   
-  dd = dist.R
-  y.within.window = y.within.window.R
-  dnorm(dd, sd=b[1])
-  
-  sse = (mean(y.within.window)- u[1])^2
-  return(sse)
-  #return(mean(y.within.window)- u[1])
+  #sse = (mean(y.within.window)- u[1])^2
+  #return(sse)
+  return(mean(y.within.window)- u[1])
 }
 
 
@@ -34,7 +22,7 @@ for (m in 1:1000) {
 (VAR = var(SE))
 (D2u[1]*b[1]/2)^2+sd1^2/(2*sqrt(pi*b[1]))
 
-(T.MSE = (D2u[126]*b/2)^2+sd1^2/(2*sqrt(pi*b)))
+(T.MSE = (D2u[1]*b[1]/2)^2+sd1^2/(2*sqrt(pi*b[1])))
 
 
 

### 2020-12-08 src/ks2.r

@@ -1,0 +1,219 @@
+## this file contains all the R functions required for a kernel
+## smoother. Do *NOT* mix simulations with functions!!
+
+## ks() is the main function implemented by Xing. Note that the main
+## purpose for writing this function is to demonstrate the underlying
+## mathematics, not for production-level use. For example, the current
+## function is based on the very inefficient for-loop in R; we almost
+## certainly will need to replace it by C/C++ in a production-level R
+## package. See ksmooth.c included in R's source code for a reference
+## implementation. Besides, it may be more efficient to implement KS
+## based on fast discrete Fourier transformation for fixed bandwidth.
+## Besides, for a production-level R package, we also need to add many
+## QC code to it, for example, it should stop if length(x) !=
+## length(y); it should have a mechanism to deal with NAs and probably
+## outliers; it also needs to check whether x is ordered or not -- KS
+## does not work for unsorted x!
+
+## x is the *ordered* vector of the covariate, y is the observed value
+## to be smoothed. bandwidth can either be a constant bandwidth or a
+## vector (adaptive bandwidth) that has the same length as x. In both
+## cases, the value (values) of bandwidth are defined as the standard
+## deviation of the Gaussian kernel. min.bandwith and max.bandwidth
+## are the minimum and maximum bandwidth. If bandwidth[i]
+## <min.bandwidth, we simply return y[j] as yhat[j]; if bandwidth[i] >
+## max.bandwidth, we use simple unweighted average over the smoothing
+## window as yhat[j]. By default ("auto"), min.bandwidth is set to be
+## min(xsteps)/4, where xsteps are the differences of two consecutive
+## x[j]; max.bandwidth is set to be 10*xrange, where xrange is the
+## range of the domain.
+
+## max.window is the maximum smoothing window, by default, it is set
+## to be 20% of xrange. For numerical reasons,
+## irrespective of the value of max.window, we don't need to extend
+## the smoothing window to more than 4*bandwidth.  Type is a
+## categorical variable that defines the type of kernel smoothing: its
+## default value "Euclidean" is the commonly used kernel smoother; the
[…]

### 2020-12-08 src/simulation.R

@@ -66,10 +66,10 @@ lines(x, yhat.fixed[, 11], col=4)
 ############################################
 
 ## specify parameters that generates data in Sim2
-N <- 501
+N <- 501   # initially 501
 ## Note that you shouldn't use "sd" as an object; it *replaces* an
 ## important system function sd()!!
-sd1 <- 1; var1 <- sd1^2
+sd1 <- 0.1; var1 <- sd1^2  # sd initially 1
 
 ## an evenly-spaced grid of x defined 
 x <- seq(-pi, pi, length.out=N); dx <- range(x)/(N-1)
@@ -137,7 +137,7 @@ legend("topleft", c("true function","Adaptive","logb=-0.6 (Opt)","logb=0"), col 
 
 #### save plots #########
 
-filename = paste('sim.plot8.eps')
+filename = paste('sim.plot11.eps')
 setEPS()
 postscript(filename,height=3,width=10)
 layout(matrix(1:3,1,3, byrow = T))
@@ -145,7 +145,7 @@ layout(matrix(1:3,1,3, byrow = T))
 
 MSE.mean <- colMeans(MSEtab); MSE.se <- apply(MSEtab, 2, sd)/sqrt(reps)
 plot(logbs, MSE.mean[-1], type="b", ylim=range(MSE.mean),
-     xlab="log-bandwidth", ylab="Mean MSE", main = "max.bandwidth = xrange/0.1")
+     xlab="log-bandwidth", ylab="Mean MSE", main = "max.bandwidth = xrange/5")
 ## +/- 2*STDERR
 lines(logbs,MSE.mean[-1]+2*MSE.se[-1], lty=3)
 lines(logbs,MSE.mean[-1]-2*MSE.se[-1], lty=3)
@@ -153,7 +153,7 @@ abline(h=MSE.mean[1], lwd=2, lty=2)
 
 
 plot(x,Res[,1], ylim = c(-1,1),type = "l", ylab = "residual",
-     main = expression(paste("Resudial Plot, ",sigma," = 1, range = [",-pi,",",pi,"]")), col = 2)
+     main = expression(paste("Resudial Plot, ",sigma," = 0.1, range = [",-pi,",",pi,"]")), col = 2)
 lines(x,Res[,6],  col = 3)
 lines(x,Res[,7],  col = 4)
 legend("top", c("Adaptive","logb=-0.6 (Opt)", "logb=0"), col =2:4, pch = 20 )

### 2021-01-19 latex/draft.tex
update the meeting notes
update code and document after reading course

@@
 \begin{itemize}
+ % 01/19/2021  
+\item 01/19/2021. Checked theoretical MSE vs empirical MSE at the whole scale. Explored the scenarios with increasing k value (to narrow the smooth.window) and increasing sample size.\\
+We noticed that the empirical MSE is smaller than theoretical MSE.
+And theoretical has certain pattern while empirical MSE is more in randomness.
+We suspect our theoretical derivation (Eq (1)) is incorrect.
+
+\begin{equation}
+         \text{MSE}_j=  \frac{\sigma^2}{2\sqrt{\pi} \cdot \sqrt{b}}  + \left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2
+\end{equation}
+
+We found the MSE of KDE online (UW lecture notes) expressed as (Eq (3))
+\begin{align}
+         \text{MSE}_j &= (nh)^{-1}f(x)\int K(z)^2 dz + \left(\frac{f''(x)}{2} h^2 \int z^2K(z)dz \right)^2 \\
+         &= \frac{u}{2\sqrt{\pi} \cdot n\sqrt{b}}  + \left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2
+\end{align}
+
+We then use their theoretical result but it still shows the certain pattern that our empirical MSE didn't reveal.
+We are wondering if our theoretical result is problematic, why our emperical result, which uses adaptive bandwidth ($b =\left(\frac{\sigma^2}{2\sqrt{\pi} u''^2}\right)^{\frac{2}{5}} $), still generates quite good estimation?
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pic/sim.diff1.eps}
+\includegraphics[width=\linewidth]{pic/sim.diff2.eps}
+\end{figure}
+
+TODOLIST:
+\begin{enumerate}
+\item rederivate formular (N should be in the denominator)
+\item plot variance and bias seperately to see which one is more problematic
+\item adaptive bandwidth should give constant MSE
+\item Think about non-constant adaptive Gaussian smoother and inhomogeneous heat equation:
+$$ \frac{\partial u}{\partial t} = \frac{\partial}{\partial x}(c \cdot \frac{\partial u}{\partial t} ) =  c\cdot u_x''$$
+when $c$ is not a constant but a function $c(x)$, the equation becomes 
+$$ \frac{\partial u}{\partial t} = c'(x)u_x' + c\cdot u_x''$$
+$c'(x)u_x' $ can be understood as tuning parameter in heat equation, where $c$ shows how fast the heat transform from one locaton to another (spatial discrepancy).
+We shall explore the relationship between $c$ in heat equation and bandwidth $b$ in kernel smoother.
+\end{enumerate}
+
   % 12/08/2020  
@@
 \begin{enumerate}
-\item 
-\item
+\item Check theoretical MSE vs empirical MSE
+$$\sin^2(t) = \sin(t) - \sin(t_0)- \cos(t_0)(t-t_0)$$
+\begin{itemize}
+\item bias$^2$ $\propto \sin^2(t)$
+\item variance $\propto \frac{\sigma^2}{n_{\texttt{neighbor}}}$
+\end{itemize}
+increase $k$ to narrow the smooth.window $ \frac{\texttt{xrange}}{k}$
+\item Map non-constant adaptive Gaussian smoother back to inhomogeneous heat equation 
 \end{enumerate}

### 2021-01-19 oldFiles/debug.R

@@ -1,28 +1,156 @@
-simk <- function(k,y){
+setwd("~/Desktop/Research/UR/HeatEquation/latex/pic")
+a=11
+
+simk <- function(k,y,a){
   smooth.window = 2*pi/k
   
-  dj <- pmin(abs(x-x[1]), x[length(x)]- x[1] - abs(x-x[1]))
+  dj <- pmin(abs(x-x[a]), x[length(x)]- x[a] - abs(x-x[a]))
   y.within.window <- y[dj <= smooth.window]
-  #dd = dj[dj <= smooth.window]
-  #w <- dnorm(dd, sd=b[1])
   
-  #sse = (mean(y.within.window)- u[1])^2
-  #return(sse)
-  return(mean(y.within.window)- u[1])
+  
+  #dd = dj[dj <= smooth.window]
+  #w <- dnorm(dd, sd=b[a])
+  
+  sse = (mean(y.within.window)- u[a])^2
+  return(sse)
+  #return(mean(y.within.window)- u[a])
 }
 
+# k=5
+# SE = NULL
+# for (m in 1:1000) {
+#   y <- u + sd1*rnorm(N)
+#   SE[m] = simk(k,y)
+# }
+# (MSE = mean(SE))
+# # (VAR = var(SE))
+# (D2u[a]*b[a]/2)^2+sd1^2/(2*sqrt(pi*b[a]))
+# 
 
-k=5
-SE = NULL
+k=25
+SE = array(dim = c(1000, length(x)))
 for (m in 1:1000) {
-  y <- u + sd1*rnorm(N)
-  SE[m] = simk(k,y)
+  for (a in 1:length(x)) {
+    y <- u + sd1*rnorm(N)
+    SE[m,a] = simk(k,y,a)
+  }
 }
-(MSE = mean(SE))
-(VAR = var(SE))
-(D2u[1]*b[1]/2)^2+sd1^2/(2*sqrt(pi*b[1]))
 
-(T.MSE = (D2u[1]*b[1]/2)^2+sd1^2/(2*sqrt(pi*b[1])))
+
+
+theoryMSE = NULL
+for (i in 1:501) {
+  theoryMSE[i] = (D2u[i]*b[i]/2)^2+sd1^2/(2*sqrt(pi*b[i]))
+}
+diff = colMeans(SE) - theoryMSE
+plot(x, diff)
+
+
+
+####################################################################################
+N <-501   # initially 501
+## Note that you shouldn't use "sd" as an object; it *replaces* an
+## important system function sd()!!
[…]

### 2021-01-19 src/ks2.r

@@ -188,7 +188,7 @@ ks.variable <- function(x, y, bandwidth, smooth.window, type=c("Euclidean", "cir
 
 
 ## this is the main wrapper
-ks <- function(x, y, bandwidth, min.bandwidth="auto", max.bandwidth="auto", max.window="auto", type=c("Euclidean", "circular")){
+ks <- function(x, y, k,bandwidth, min.bandwidth="auto", max.bandwidth="auto", max.window="auto", type=c("Euclidean", "circular")){
   type <- match.arg(type)
   N <- length(x); xsteps <- diff(x); xrange <- max(x)-min(x)
   if (min.bandwidth=="auto") min.bandwidth <- min(xsteps)

### 2021-01-19 src/simulation.R

@@ -74,9 +74,9 @@ sd1 <- 0.1; var1 <- sd1^2  # sd initially 1
 ## an evenly-spaced grid of x defined 
 x <- seq(-pi, pi, length.out=N); dx <- range(x)/(N-1)
 ## the true function to be estimated
-u <- sin(x)
+u <- sin(x)+cos(x)
 ## its second order derivative
-D2u <- -sin(x)
+D2u <- -sin(x)-cos(x)
 ## the optimal bandwidth
 b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
 

### 2021-02-04 latex/draft.tex
update files after meeting

@@
 \usepackage{float}
+\usepackage{color}
 
@@
 \begin{itemize}
+
+ % 02/02/2021  
+\item 02/02/2021.
+
+We fixed the theoretical formula for both variance and bias:
+\begin{equation}
+         \text{MSE}_j=   {\color{red}\frac{L}{N-1}}\cdot \frac{\sigma^2}{2\sqrt{\pi} \cdot {\color{red}b}}  + \left[\frac{\partial^2 u}{2\partial x^2}\cdot b {\color{red}^2} \right]^2
+\end{equation}
+
+We plot the variance and bias$^2$ scenarios separately.
+
+\begin{figure}[H]
+Plot sample size $N$ versus variance (left is the empirical variance mean, and right is the certain point variance.)\\
+\includegraphics[width=2.5in]{pic/variance.png}
+\includegraphics[width=2.5in]{pic/variance2.png}
+\end{figure}
+
+\begin{figure}[H]
+Plot smooth.window parameter $K$ versus variance\\
+\includegraphics[width=2.5in]{pic/K.png}
+\includegraphics[width=2.5in]{pic/K2.png}
+\end{figure}
+
+\begin{figure}[H]
+Plot sample size $N$ versus variance\\
+\includegraphics[width=1.55in]{pic/bias1.png}
+\includegraphics[width=1.55in]{pic/bias2.png}
+\includegraphics[width=1.55in]{pic/bias3.png}
+\end{figure}
+
+\begin{figure}[H]
+Plot sample size $N$ versus variance\\
+\includegraphics[width=2.5in]{pic/bias4.png}
+\includegraphics[width=2.5in]{pic/bias5.png}
+\end{figure}
+
+ % 01/26/2021  
+\item 01/26/2021.
+
+In order to find out the reason that empirical MSE and theoretical MSE look so different, we split them into $\texttt{bias}^2$ and \texttt{variance}.
+
+For fixed bandwidth scenario, we tried with the below 3 possible theoretical MSE with sample size $N=501$ and simulation time $n=500$:
+
+\begin{equation}
+         \text{MSE}_j=  \frac{\sigma^2}{2\sqrt{\pi} \cdot \sqrt{b}}  + \left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2
+\end{equation}
+
+\begin{equation}
+         \text{MSE}_j=  \frac{\sigma^2}{2N\sqrt{\pi} \cdot \sqrt{b}}  + \frac{20}{N^2}\cdot\left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2
+\end{equation}
+
+\begin{equation}
+         \text{MSE}_j=  \frac{\sigma^2}{2\sqrt{N}\sqrt{\pi} \cdot \sqrt{b}}  + \frac{20}{N}\cdot\left[\frac{\partial^2 u}{2\partial x^2}\cdot b \right]^2
+\end{equation}
+
+Empirical MSE:
+\begin{align*}
+         \text{MSE}_j&=  \mathbb{E}[(\hat{y} - u)^2]\\
+         &= \mathbb{E}[(\hat{y} - \mathbb{E}(\hat{y}))^2] + \mathbb{E}[\mathbb{E} (\hat{y}) -u]^2\\ 
+          &= \texttt{var}(\hat{y}) + \texttt{bias}^2(\hat{y})
+\end{align*}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pic/sim.mse.eps}
+\includegraphics[width=\linewidth]{pic/sim.mse2.eps}
+Increase sample size from N=501 to N=1001:\\
+\includegraphics[width=\linewidth]{pic/sim.mse3.eps}
+\includegraphics[width=\linewidth]{pic/sim.mse4.eps}
+\end{figure}
+
+TODOLIST:
+\begin{enumerate}
+\item plot \texttt{variance} vs sample size (N) plot to see the relationship
+\item take a close look of \texttt{bias} term: Taylor expansion on a fix point (eg.: $f(x) = \sin(x)$), to see if the first term goes to $f(x_0)$, second term goes to 0 and third term goes to $f''(x_0)b^2/2$
+\item After solve the fix bandwidth scenario issue, try adaptive bandwidth
+\end{enumerate}
+
  % 01/19/2021  

### 2021-02-04 oldFiles/debug.R

@@ -48,7 +48,7 @@ plot(x, diff)
 
 
 ####################################################################################
-N <-501   # initially 501
+N <-1001   # initially 501
 ## Note that you shouldn't use "sd" as an object; it *replaces* an
 ## important system function sd()!!
 sd1 <- 0.1; var1 <- sd1^2  # sd initially 1
@@ -60,42 +60,49 @@ u <- sin(x)
 ## its second order derivative
 D2u <- -sin(x)
 ## the optimal bandwidth
-b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+#b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+b = rep(0.2, N)
 
-
-simk <- function(y,x,u,k){
+simk <- function(y,x,u,k,b){
   yhat = ks(x, y, k,bandwidth=b, type = "Euclidean")
   sse = (yhat - u)^2
   return(sse)
 }
 
-simk2 <- function(y,x,u,k){
+simk2 <- function(y,x,u,k,b){
   yhat = ks(x, y, k,bandwidth=b, type = "circular")
   sse = (yhat - u)^2
   return(sse)
 }
 
 
-SE = array(dim = c(500, length(x)))
-for (m in 1:500) {
+SE = array(dim = c(50, length(x)))
+for (m in 1:50) {
     y <- u + sd1*rnorm(N)
-    SE[m,] = simk(y,x,u,5)
+    SE[m,] = simk(y,x,u,5,b)
     print(m)
 }
 
 SE2 = array(dim = c(500, length(x)))
+y = array(dim = c(500, length(x)))
+yhat = array(dim = c(500, length(x)))
 for (m in 1:500) {
-  y <- u + sd1*rnorm(N)
-  SE2[m,] = simk2(y,x,u,5)
+  y[m,] <- u + sd1*rnorm(N)
+  #SE2[m,] = simk2(y,x,u,5,b)
+  yhat[m,] = ks(x, y[m,], bandwidth=b, type = "circular")
   print(m)
 }
 
+variance = colMeans((yhat-matrix(rep(colMeans(yhat),500), nrow = 500, byrow = T ))^2)
+biass = (colMeans(yhat)-u)^2
 
 theoryMSE = NULL
 theoryMSE2 = NULL
-for (i in 1:501) {
-  theoryMSE[i] = (D2u[i]*b[i]/2)^2+ sd1^2/(2*sqrt(pi*b[i]))
[…]

### 2021-02-04 oldFiles/pde.bias.r

@@ -1,0 +1,42 @@
+N=1001
+sd1 <- 0.5; var1 <- sd1^2  # sd initially 1
+## an evenly-spaced grid of x defined 
+x <- seq(-pi, pi, length.out=N); dx <- range(x)/(N-1)
+L = diff(range(x))
+## the true function to be estimated
+u <- x^4
+## its second order derivative
+D2u <- 12*x^2
+## the optimal bandwidth
+#b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+b = 0.2
+
+simk2 <- function(y,x,u,k,b){
+  yhat = ks(x, y, k,bandwidth=b, type = "circular")
+  sse = (yhat - u)^2
+  return(sse)
+}
+
+set.seed(123)
+y = array(dim = c(100, length(x)))
+yhat = array(dim = c(100, length(x)))
+for (m in 1:100) {
+  y[m,] <- u + sd1*rnorm(N)
+  #SE2[m,] = simk2(y,x,u,5,b)
+  yhat[m,] = ks(x, y[m,], k=5, bandwidth=b, type = "circular")
+  print(m)
+}
+
+biass = (colMeans(yhat)-u)^2
+
+
+
+theoryMSE = NULL
+for (i in 1:N) {
+  theoryMSE[i] = (0.5*D2u[i]*b^2)^2
+}
+
+plot(x, theoryMSE, main="Bias (N=1001, K=5), y=x^4",col = 1,pch=19, xlim = c(-2,2),ylim = c(0,1))
+points(x, biass, col=2, pch=20 )
+legend("top", c("theory", "emperical"), pch=c(19,20),col=1:2, bty="n")
+

### 2021-02-04 oldFiles/pde.variance.r

@@ -1,0 +1,63 @@
+# Variance with sigma
+
+
+#samplesize <-c(251,501,751,1001)  # initially 501
+
+## Note that you shouldn't use "sd" as an object; it *replaces* an
+## important system function sd()!!
+K = c(0.1,5,8,15)
+theory.N = variance.N = array(dim = 4)
+for (j in 1:4) {
+  #N = samplesize[j]
+  N=1001
+  sd1 <- 0.5; var1 <- sd1^2  # sd initially 1
+  ## an evenly-spaced grid of x defined 
+  x <- seq(-pi, pi, length.out=N); dx <- range(x)/(N-1)
+  L = diff(range(x))
+  ## the true function to be estimated
+  u <- sin(x)
+  ## its second order derivative
+  D2u <- -sin(x)
+  ## the optimal bandwidth
+  #b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+  b = 0.2
+  
+  simk2 <- function(y,x,u,k,b){
+    yhat = ks(x, y, k,bandwidth=b, type = "circular")
+    sse = (yhat - u)^2
+    return(sse)
+  }
+  
+  set.seed(123)
+  y = array(dim = c(500, length(x)))
+  yhat = array(dim = c(500, length(x)))
+  for (m in 1:500) {
+    y[m,] <- u + sd1*rnorm(N)
+    #SE2[m,] = simk2(y,x,u,5,b)
+    yhat[m,] = ks(x, y[m,], k=K[j], bandwidth=b, type = "circular")
+    print(m)
+  }
+  
+  variance = colMeans((yhat-matrix(rep(colMeans(yhat),500), nrow = 500, byrow = T ))^2)
+
+  theoryMSE = NULL
+  for (i in 1:N) {
+    #theoryMSE[i] = sd1^2*L/((N-1)*sqrt(pi*b))
+    theoryMSE[i] = sd1^2*L/((N-1)*2*sqrt(pi)*b)
+  }
+  
+  theory.N[j] = mean(theoryMSE)
+  variance.N[j] <- mean(variance)
+  #variance.N[j] <- variance[30]
+}
+
+
+
+plot(K, theory.N, pch=20, ylim = c(0, max(theory.N, variance.N)+0.001), main="Variance", col = 1)
+points(K, variance.N, col=2, pch=20 )
+legend("topright", c("theory", "emperical"), pch=20,col=1:2)
+
+
+
+temp = cbind(samplesize, theory.N, variance.N)
+

### 2021-02-04 src/ks2.r

@@ -193,7 +193,7 @@ ks <- function(x, y, k,bandwidth, min.bandwidth="auto", max.bandwidth="auto", ma
   N <- length(x); xsteps <- diff(x); xrange <- max(x)-min(x)
   if (min.bandwidth=="auto") min.bandwidth <- min(xsteps)
   if (max.bandwidth=="auto") max.bandwidth <- xrange*10  # initially 10*xrange
-  if (max.window=="auto") max.window <- xrange/5  # initially xrange/5
+  if (max.window=="auto") max.window <- xrange/k # initially xrange/5
   ## classify the smoothing strategies based on bandwidth
   if (length(bandwidth)==1){ #fixed bandwidth
     if (bandwidth <= min.bandwidth) {

### 2021-02-16 latex/draft.tex
update after meeting

@@
 \begin{itemize}
+
+ % 02/08/2021  
+\item 02/08/2021. 
+\begin{equation}
+         b = \big( \frac{L\sigma^2}{2\sqrt{\pi}(N-1)f''}  \big)^{1/5}
+\end{equation}
+
+\begin{figure}[H]
+Plot MSE with logbandwidth (left: $y = x^4$; right: $y = \sin(x)$)\\
+\includegraphics[width=2.5in]{pic/mse.bw0.png}
+\includegraphics[width=2.5in]{pic/mse.bw.png}
+\end{figure}
+
+\begin{figure}[H]
+Plot MSE with logbandwidth with confidence interval \\($y = \sin(x)$, N = 2001, rep = 200)\\
+\includegraphics[width=2.5in]{pic/mse.ci0.png}
+\includegraphics[width=2.5in]{pic/mse.ci.png}
+\end{figure}
+
+\begin{figure}[H]
+Plot sample size $N$ versus bias (using adaptive bandwidth)\\
+\includegraphics[width=1.55in]{pic/bias6.png}
+\includegraphics[width=1.55in]{pic/bias7.png}
+\includegraphics[width=1.55in]{pic/bias8.png}
+\end{figure}
+
+\begin{figure}[H]
+We notice that when bandwidth is too small or too large, the values fall out of the smooth window (2 left figures: small fixed b; right figure: large fixed b)\\
+\includegraphics[width=1.55in]{pic/bias9.1.png}
+\includegraphics[width=1.55in]{pic/bias9.2.png}
+\includegraphics[width=1.55in]{pic/bias10.png}
+\end{figure}
 
@@
 \begin{figure}[H]
-Plot sample size $N$ versus variance\\
+Plot sample size $N$ versus bias\\
 \includegraphics[width=1.55in]{pic/bias1.png}
@@
 \begin{figure}[H]
-Plot sample size $N$ versus variance\\
+Plot sample size $N$ versus bias\\
 \includegraphics[width=2.5in]{pic/bias4.png}
@@
 \end{figure}
+
+TODOLIST:
+\begin{enumerate}
+\item find the optimal smoother and redraw the figures. 
+\item prove adaptive smoothing is an advantage
+\item replace variate kernel smoother to heat equation
+\begin{align*}
+\frac{\partial u}{\partial t} &= \nabla_x (c(x)\nabla_x u) \\ 
+&\approx c(x) \nabla^2 u \\ 
+\end{align*}
+\item implement finite differential solver (compare diffusion coefficient)
+\item long term: finite different method (FDM) and finite element method (FEM) 
+\end{enumerate}
 

### 2021-02-16 oldFiles/pde.bias.r

@@ -4,12 +4,18 @@ sd1 <- 0.5; var1 <- sd1^2  # sd initially 1
 x <- seq(-pi, pi, length.out=N); dx <- range(x)/(N-1)
 L = diff(range(x))
 ## the true function to be estimated
-u <- x^4
+#u <- sin(x/5)*25
+u= sin(x)
+#u <- x^4
 ## its second order derivative
-D2u <- 12*x^2
+#D2u <- -sin(x/5)
+D2u <- -sin(x)
+#D2u <- 12*x^2
 ## the optimal bandwidth
-#b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+b = (L*var1/(2*sqrt(pi)*(N-1)*D2u^2))^(1/5)
+b = 1
 b = 0.2
+b = 0.001
 
 simk2 <- function(y,x,u,k,b){
   yhat = ks(x, y, k,bandwidth=b, type = "circular")
@@ -18,9 +24,9 @@ simk2 <- function(y,x,u,k,b){
 }
 
 set.seed(123)
-y = array(dim = c(100, length(x)))
-yhat = array(dim = c(100, length(x)))
-for (m in 1:100) {
+y = array(dim = c(500, length(x)))
+yhat = array(dim = c(500, length(x)))
+for (m in 1:500) {
   y[m,] <- u + sd1*rnorm(N)
   #SE2[m,] = simk2(y,x,u,5,b)
   yhat[m,] = ks(x, y[m,], k=5, bandwidth=b, type = "circular")
@@ -36,7 +42,7 @@ for (i in 1:N) {
   theoryMSE[i] = (0.5*D2u[i]*b^2)^2
 }
 
-plot(x, theoryMSE, main="Bias (N=1001, K=5), y=x^4",col = 1,pch=19, xlim = c(-2,2),ylim = c(0,1))
-points(x, biass, col=2, pch=20 )
+plot(x, theoryMSE, main="Bias, y=sin(x), (b=1)",col = 1,pch=19)
+points(x, biass, col=2, pch=20,main="Bias, y=sin(x), (b=0.001)" )
 legend("top", c("theory", "emperical"), pch=c(19,20),col=1:2, bty="n")
 

### 2021-02-16 src/ks2.r

@@ -188,7 +188,7 @@ ks.variable <- function(x, y, bandwidth, smooth.window, type=c("Euclidean", "cir
 
 
 ## this is the main wrapper
-ks <- function(x, y, k,bandwidth, min.bandwidth="auto", max.bandwidth="auto", max.window="auto", type=c("Euclidean", "circular")){
+ks <- function(x, y, k=5,bandwidth, min.bandwidth="auto", max.bandwidth="auto", max.window="auto", type=c("Euclidean", "circular")){
   type <- match.arg(type)
   N <- length(x); xsteps <- diff(x); xrange <- max(x)-min(x)
   if (min.bandwidth=="auto") min.bandwidth <- min(xsteps)

### 2021-02-16 src/simulation.R

@@ -10,12 +10,13 @@ sd1 <- 50; var1 <- sd1^2
 
 ## an evenly-spaced grid of x defined on [-5,5]
 x <- seq(-5, 5, length.out=N); dx <- 10/(N-1)
+L = diff(range(x))
 ## the true function to be estimated
 u <- x^4+600
 ## its second order derivative
 D2u <- 12*x^2
 ## the optimal bandwidth
-b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+b = (L*var1/(2*sqrt(pi)*(N-1)*D2u^2))^(1/5)
 
 ## the fixed bandwidths
 min.logb <- -1; max.logb <- -0.7; nb <- 11
@@ -72,23 +73,24 @@ N <- 501   # initially 501
 sd1 <- 0.1; var1 <- sd1^2  # sd initially 1
 
 ## an evenly-spaced grid of x defined 
-x <- seq(-pi, pi, length.out=N); dx <- range(x)/(N-1)
+x <- seq(-pi/2, 0, length.out=N); dx <- range(x)/(N-1)
+L = diff(range(x))
 ## the true function to be estimated
-u <- sin(x)+cos(x)
+u <- sin(x)
 ## its second order derivative
-D2u <- -sin(x)-cos(x)
+D2u <- -sin(x)
 ## the optimal bandwidth
-b = (var1/(2*sqrt(pi)*D2u^2))^(2/5)
+b = (L*var1/(2*sqrt(pi)*(N-1)*D2u^2))^(1/5)
 
 
-plot(x, b, type="l", ylim=c(0,5), xlab="x", ylab="Adaptive bandwidth")
+plot(x, b, type="l",  xlab="x", ylab="Adaptive bandwidth")
 
 ## the fixed bandwidths
-min.logb <- -3; max.logb <- 3; nb <- 11
+min.logb <- -1; max.logb <- 0; nb <- 11
 logbs <- seq(min.logb, max.logb, length.out=nb)
 
 ## number of repetitions
-reps <- 50
+reps <- 100
 
 # Relationship between the smoothing bandwidth and mean MSE
 MSEtab <- matrix(0, nrow=reps, ncol=nb+1)
@@ -97,17 +99,25 @@ set.seed(123)
 for (j in 1:reps){
   print(paste0("Repetition: ", j))
   y <- u + sd1*rnorm(N)
[…]
